---
title: "Vertical Habitat Use of NA Blue and Mako Sharks"
author: "Aidan Cox"
date: "1/5/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# LIBRARY LIST

``` {r Library List, echo=FALSE}
# general #
library(tidyverse)
library(ggplot2)
library(writexl)
library(readxl)

# clustering #
library(NbClust)
library(tags2etuff)
library(devtools)
library(dendextend)
library(hms)
library(vegan)
library(ggdendro)

# mapping #
library(rnaturalearth)
library(rnaturalearthdata)
library(rgeos)
library(ggspatial)
library(sf)
library(viridis)
library(ncdf4)

# environmental data #
library(readr)
library(raster)
library(HMMoce)
library(lunar)
devtools::load_all('~/Desktop/blues-makos/analyzePSAT')

# logistic regression #
library(foreign)
library(nnet)
library(reshape2)
library(mclogit)

world <- ne_countries(scale = "medium", returnclass = "sf")
```

## DATA PROCESSING
# Load the etuff data for blue and mako sharks ---- 
``` {r Read etuff files}
fList <- list.files("~/Desktop/blues-makos/etuff_new/", full.names = T)
blues <- fList[grep('160424', fList)]
makos <- fList[grep('159924', fList)]

# for mako sharks
for(i in 1:length(makos)) {
  etuff <- read_etuff(makos[i])
  tad <- get_tad(etuff)
  tad <- tad %>% separate(DateTime, into = c("Date", "Time"), sep = "([ ])")
  tad <- tad %>% pivot_wider(id_cols = Date,
              names_from = bin,
              values_from = freq,
              values_fn = list(freq = mean))
  tad$ptt <- etuff$meta$ptt
  tad$owner <- etuff$meta$person_owner
  assign(paste("mtad", i, sep = "_"), tad)
  rm(etuff, tad)
  print(i)
}

# for blue sharks 
for(i in 1:length(blues)) {
  etuff <- read_etuff(blues[i])
  tad <- get_tad(etuff)
  tad <- tad %>% separate(DateTime, into = c("Date", "Time"), sep = "([ ])")
  tad <- tad %>% pivot_wider(id_cols = Date,
                             names_from = bin,
                             values_from = freq,
                             values_fn = list(freq = mean))
  tad$ptt <- etuff$meta$ptt
  tad$owner <- etuff$meta$person_owner
  assign(paste("btad", i, sep = "_"), tad)
  rm(etuff, tad)
  print(i)
}
```

# now establish common bins between the different tags
``` {r Common Depth Bins}
# first make sure that Greg's tags have 14 tad bins and Cam's tags have 12
# makos
mtad_3$TimeAtDepthBin12 <- 0 # Greg's tag
mtad_3$TimeAtDepthBin13 <- 0 # Greg's tag
mtad_3$TimeAtDepthBin14 <- 0 # Greg's tag

mtad_5$TimeAtDepthBin11 <- 0 # Cam's tag
mtad_5$TimeAtDepthBin12 <- 0 # Cam's tag

mtad_6$TimeAtDepthBin12 <- 0 # Cam's tag

mtad_7$TimeAtDepthBin12 <- 0 # Cam's tag

# blues (all Cam's tags)
btad_1$TimeAtDepthBin11 <- 0
btad_1$TimeAtDepthBin12 <- 0

btad_2$TimeAtDepthBin12 <- 0

btad_3$TimeAtDepthBin12 <- 0

btad_7$TimeAtDepthBin12 <- 0

btad_9$TimeAtDepthBin11 <- 0
btad_9$TimeAtDepthBin12 <- 0

btad_11$TimeAtDepthBin12 <- 0

# now combine bins to create the below distribution
bins_max <- c(10, 50, 100, 200, 300, 400, 500, 2000)
# Camrin's Innate Bins = c(2, 10, 20, 50, 100, 200, 300, 400, 500, 700, 1000, 2000)
# Gregory's Innate Bins = c(5, 10, 25, 50, 75, 100, 150, 200, 300, 400, 500, 600, 800, 2000)

# makos
gregory_mako <- rbind(mtad_1, mtad_2, mtad_3) 
gregory_mako <- gregory_mako %>% transmute(
  Date = Date,
  bin1 = TimeAtDepthBin01 + TimeAtDepthBin02,
  bin2 = TimeAtDepthBin03 + TimeAtDepthBin04,
  bin3 = TimeAtDepthBin05 + TimeAtDepthBin06,
  bin4 = TimeAtDepthBin07 + TimeAtDepthBin08,
  bin5 = TimeAtDepthBin09,
  bin6 = TimeAtDepthBin10,
  bin7 = TimeAtDepthBin11,
  bin8 = TimeAtDepthBin12 + TimeAtDepthBin13 + TimeAtDepthBin14,
  ptt = ptt,
  owner = owner
)
camrin_mako <- rbind(mtad_4, mtad_5, mtad_6, mtad_7)
camrin_mako <- camrin_mako %>% transmute(
  Date = Date,
  bin1 = TimeAtDepthBin01 + TimeAtDepthBin02,
  bin2 = TimeAtDepthBin03 + TimeAtDepthBin04,
  bin3 = TimeAtDepthBin05,
  bin4 = TimeAtDepthBin06,
  bin5 = TimeAtDepthBin07,
  bin6 = TimeAtDepthBin08,
  bin7 = TimeAtDepthBin09,
  bin8 = TimeAtDepthBin10 + TimeAtDepthBin11 + TimeAtDepthBin12,
  ptt = ptt,
  owner = owner
)

tad_mako <- rbind(gregory_mako, camrin_mako)

rm(gregory_mako, camrin_mako)

# blues 

tad_blues <- rbind(btad_1, btad_2, btad_3, btad_4, btad_5, btad_6, btad_7, btad_8, btad_9, btad_10, btad_11, btad_12, btad_13)
tad_blues <- tad_blues %>% transmute(
  Date = Date,
  bin1 = TimeAtDepthBin01 + TimeAtDepthBin02,
  bin2 = TimeAtDepthBin03 + TimeAtDepthBin04,
  bin3 = TimeAtDepthBin05,
  bin4 = TimeAtDepthBin06,
  bin5 = TimeAtDepthBin07,
  bin6 = TimeAtDepthBin08,
  bin7 = TimeAtDepthBin09,
  bin8 = TimeAtDepthBin10 + TimeAtDepthBin11 + TimeAtDepthBin12,
  ptt = ptt,
  owner = owner
)
```

```{r cleanup}
rm(mtad_1, mtad_2, mtad_3, mtad_4, mtad_5, mtad_6, btad_1, btad_2, btad_3, btad_4, btad_5, btad_6, btad_7, btad_8, btad_9, btad_10, btad_11, btad_12, btad_13 )
```

## GENERAL COMPARISONS
# Time-at-temperature (general)
Construct time at temperature profiles by adding temperature data to time-series of depth occupancy. Note: For mako sharks, only camrin's tags can be used for this computation since greg's tags don't contian series information

``` {r Time-at-temperature}
# MAKOS
pdt <- data.frame()

for (i in 4:length(makos)) {
  etuff <- read_etuff(makos[i])
  hdr <- get_header(makos[i])
  temp <- get_pdt(etuff)
  interp <- interp_pdt(etuff)
  series <- get_series(etuff)
  series <- filter(series, !is.na(depth))
  series <- add_series_temp(series, temp, interp)
  series <- filter(series, !is.na(temperature))
  series$dn <- add_daynight(series, etuff)
  series$ptt <- hdr$ptt
  pdt <- rbind(pdt, series)
  rm(etuff, hdr, temp, interp, series)
  print(i)
}

ggplot(data = pdt) +
  geom_histogram(data = subset(pdt, dn == "d"), aes(x = temperature, y = -(..count..)/sum(..count..)), fill = NA, color = "black") +
  geom_histogram(data = subset(pdt, dn =="n"), aes(x = temperature, y = (..count..)/sum(..count..)), color = "black") +
  coord_flip() +
  ylab("Proportion") +
  xlab("Temperature (ºC)") +
  ggtitle("Mako Shark Time at Temperature") +
  theme_classic()

# BLUES
pdtb <- data_frame()

for (i in 1:length(blues)) {
  etuff <- read_etuff(blues[i])
  hdr <- get_header(blues[i])
  temp <- get_pdt(etuff)
  interp <- interp_pdt(etuff)
  series <- get_series(etuff)
  series <- filter(series, !is.na(depth))
  series <- add_series_temp(series, temp, interp)
  series <- filter(series, !is.na(temperature))
  series$dn <- add_daynight(series, etuff)
  series$ppt <- hdr$ptt
  pdtb <- rbind(pdtb, series)
  rm(etuff, hdr, temp, interp, series)
  print(i)
}

ggplot() +
  geom_histogram(data = subset(pdtb, dn == "d"), aes(x = temperature, y = -(..count..)/sum(..count..)), fill = NA, color = "black") +
  geom_histogram(data = subset(pdtb, dn == "n"), aes(x = temperature, y = (..count..)/sum(..count..)), color = "black") +
  coord_flip() +
  ylab("Proportion") +
  xlab("Temperature (ºC)") +
  ggtitle("Blue Shark Time at Temperature") +
  theme_classic()
```

# Time-at-depth (general)
``` {r Time-at-depth}
# MAKOS
ggplot() +
  geom_histogram(data = subset(pdt, dn == "d"), aes(x = depth, y = -(..count..)/sum(..count..)), fill = NA, color = "black") +
  geom_histogram(data = subset(pdt, dn == "n"), aes(x = depth, y = (..count..)/sum(..count..)), color = "black") +
  scale_x_reverse() +
  scale_y_continuous(limits = c(-0.55, 0.55), breaks = c(-0.4, -0.2, 0, 0.2, 0.4), labels = c("0.4", "0.2", "0", "0.2", "0.4")) +
  coord_flip() +
  ylab("Proportion") +
  xlab("Depth (m)") +
  ggtitle("Mako Time at Depth") +
  theme_classic()

# BLUES

ggplot() +
  geom_histogram(data = subset(pdtb, dn == "d"), aes(x = depth, y = -(..count..)/sum(..count..)), fill = NA, color = "black") +
  geom_histogram(data = subset(pdtb, dn == "n"), aes(x = depth, y = (..count..)/sum(..count..)), color = "black") +
  expand_limits(x = c(0, 1300)) +
  scale_x_reverse(breaks = c(0, 400, 800, 1200), labels = c("0", "400", "800", "1200")) +
  scale_y_continuous(breaks = c(-0.4, -0.2, 0, 0.2, 0.4), labels = c("0.4", "0.2", "0", "0.2", "0.4")) +
  coord_flip() +
  ylab("Proportion") +
  xlab("Depth (m)") +
  ggtitle("Blue Time at Depth") +
  theme_classic()

# COMBO
# Note: the binwidth here is 50, but it occurs around breaks in x (ie. -25-25, 25-75, etc.)
ggplot() +
  geom_histogram(data = pdtb, aes(x = depth, y = (..count..)/sum(..count..)), fill = NA, color = "blue", binwidth = 50) +
  geom_histogram(data = pdt, aes(x = depth, y = -(..count..)/sum(..count..)), color = "black", binwidth = 50) +
  expand_limits(x = c(0, 1300)) +
  scale_x_reverse(breaks = c(0, 400, 800, 1200, 1300), labels = c("0", "400", "800", "1200", "1300")) +
  scale_y_continuous(breaks = c(-0.25, 0, 0.25, 0.5), labels = c("0.25", "0", "0.25", "0.5")) +
  coord_flip() +
  ylab("Proportion of Time at Depth") +
  xlab("Depth (m)") +
  ggtitle("General Vertical Habitat") +
  theme_classic()

```

# Proportion Time-at-depth heatmap
Calculate the liklihood of depth occupancy across a 24-hour period. 

NOTE: Consider these figures carefully - are the same number of observations being counted at each time of day. Why is the normalized density being multiplied by the number of observations?

``` {r TAD Heatmaps}
# MAKOS
props <- pdt %>% separate(DateTime_local, c("Date", "Time_Local"), sep =  "([ ])") %>%
  group_by(Time_Local) %>%
  count(depth)

duptimes <- props$n
idx <- rep(1:nrow(props), duptimes)
duprops <- props[idx,]
duprops <- duprops[,-c(3)]
rm(props,duptimes, idx)

duprops$Time_Local <- as.hms(duprops$Time_Local)
duprops$Time_Local <- as.numeric(duprops$Time_Local)

ggplot(data = duprops, aes(x = Time_Local, y = depth) ) +
  stat_density_2d(aes(fill = ((..density..)*n)), geom = "raster", contour = FALSE) +
  scale_fill_distiller(palette = "Spectral") + 
  scale_y_reverse() +
  scale_x_continuous(breaks = c(0, 43200, 86400), labels = c("00:00:00", "12:00:00", "23:00:00")) +
  xlab("Time of Day") +
  ylab("Depth(m)") +
  theme_minimal()

# BLUES

propsb <- pdtb %>% separate(DateTime_local, c("Date", "Time_Local"), sep =  "([ ])") %>%
  group_by(Time_Local) %>%
  count(depth)

duptimes <- propsb$n
idx <- rep(1:nrow(propsb), duptimes)
dupropsb <- propsb[idx,]
dupropsb <- dupropsb[,-c(3)]
rm(propsb, duptimes, idx)

dupropsb$Time_Local <- as.hms(dupropsb$Time_Local)
dupropsb$Time_Local <- as.numeric(dupropsb$Time_Local)

ggplot(data = dupropsb, aes(x = Time_Local, y = depth) ) +
  stat_density_2d(aes(fill = ((..density..)*n)), geom = "raster", contour = FALSE) +
  scale_fill_distiller(palette = "Spectral") + 
  scale_y_reverse() +
  scale_x_continuous(breaks = c(0, 43200, 86400), labels = c("00:00:00", "12:00:00", "23:00:00")) +
  xlab("Time of Day") +
  ylab("Depth(m)") +
  theme_minimal()
```

# Proportion Time-at-temperature heatmap
Calculate the liklihood of temperature occupancy across a 24-hour period:

NOTE: Beware of a similar connundrum as in the above chunk...
``` {r TAT Heatmap}
# makos
props <- pdt %>% separate(DateTime_local, c("Date", "Time_Local"), sep =  "([ ])") %>%
  group_by(Time_Local) %>%
  count(temperature)

ggplot(data = props, aes(x = as.hms(Time_Local), y = temperature) ) +
  stat_density_2d(aes(fill = ((..density..)*n)), geom = "raster", contour = FALSE) +
  scale_fill_distiller(palette = "Spectral") + 
  scale_x_continuous(breaks = c(0, 43200, 86400), labels = c("00:00:00", "12:00:00", "23:00:00")) +
  xlab("Time of Day") +
  ylab("Temperature(ºC)") +
  theme_minimal()

# blues
propsb <- pdtb %>% separate(DateTime_local, c("Date", "Time_Local"), sep =  "([ ])") %>%
  group_by(Time_Local) %>%
  count(temperature)

ggplot(data = propsb, aes(x = as.hms(Time_Local), y = temperature) ) +
  stat_density_2d(aes(fill = ((..density..)*n)), geom = "raster", contour = FALSE) +
  scale_fill_distiller(palette = "Spectral") + 
  scale_x_continuous(breaks = c(0, 43200, 86400), labels = c("00:00:00", "12:00:00", "23:00:00")) +
  xlab("Time of Day") +
  ylab("Temperature(ºC)") +
  theme_minimal()

# cleanup
rm(props, propsb, duptimes, idx, duprops, dupropsb)
```
# Time-series information
These loops compile the time-series data from each of Camrin's tagged sharks (Greg's tags did not collect this information)

```{r get_series}
# makos
for (i in 4:length(makos)) {
  etuff <- read_etuff(makos[i])
  series <- get_series(etuff)
  series$dn <- add_daynight(series, etuff)
  series$ptt <- etuff$meta$ptt
  series$dup <- series$DateTime_local
  series <- series %>% separate(dup, into = c("Date", "Local_Time"), sep = "([ ])")
  series <- series %>% mutate(
    kode = paste(Date, ptt, sep = "_")
  )
  series <- series %>% filter(!is.na(depth))
  series$species <- "I.oxyrinchus"
  assign(paste("mseries",i, sep = "_" ), series)
  rm(etuff, series)
  print(i)
}

mako_series <- rbind(mseries_4, mseries_5, mseries_6, mseries_7)

# blues
for (i in 1:length(blues)) {
  etuff <- read_etuff(blues[i])
  series <- get_series(etuff)
  series$dn <- add_daynight(series, etuff)
  series$ptt <- etuff$meta$ptt
  series$dup <- series$DateTime_local
  series <- series %>% separate(dup, into = c("Date", "Local_Time"), sep = "([ ])")
  series <- series %>% mutate(
    kode = paste(Date, ptt, sep = "_")
  )
  series <- series %>% filter(!is.na(depth))
  series$species <- "P.glauca"
  assign(paste("bseries",i, sep = "_" ), series)
  rm(etuff, series)
  print(i)
}

bseries_12 <- bseries_12[,-4]
blue_series <- rbind(bseries_1, bseries_2, bseries_3, bseries_4, bseries_5, bseries_6, bseries_7, bseries_8, bseries_9, bseries_10, bseries_11, bseries_12, bseries_13)
```

# general vertical habitat figure
Here we will compile series data from all tagged sharks which recorded data from 2016-01-18 through 2016-01-25:
``` {r series general vert. habitat}
bs <- blue_series %>% filter(Date == "2016-01-18" | Date == "2016-01-19" | Date == "2016-01-20" | Date == "2016-01-21" | Date == "2016-01-22" | Date == "2016-01-23" | Date == "2016-01-24" | Date == "2016-01-25")
bs <- bs %>% filter(ptt == 141259)
bs$DateTime_local <- bs$DateTime - 18000
bs$species <- "P.glauca"

ms <- mako_series %>% filter(Date == "2016-01-18" | Date == "2016-01-19" | Date == "2016-01-20" | Date == "2016-01-21" | Date == "2016-01-22" | Date == "2016-01-23" | Date == "2016-01-24" | Date == "2016-01-25")
ms$DateTime_local <- ms$DateTime - 18000
ms$species <- "I.oxyrinchus"

cs <- rbind(ms, bs)
cs$ptt <- factor(cs$ptt)

ggplot() +
  geom_line(data = cs, aes(x = DateTime_local, y = depth, color = species)) +
  scale_y_reverse() +
  facet_grid(species~.) +
  theme_classic()
  
```

# Tagging location map
``` {r Tagging locations and track maps}
# makos
for(i in 1:length(makos)) {
  etuff <- read_etuff(makos[i])
  track <- get_track(etuff)
  track$ptt <- etuff$meta$ptt
  track <- track %>% separate(DateTime, into = c("Date", "Time"), sep = "([ ])")
  track <- track %>% mutate(
    kode = paste(Date, ptt, sep = "_")
  )
  assign(paste("mtrack",i, sep = "_" ), track)
  rm(etuff, track)
  print(i)
}

mako_tracks <- rbind(mtrack_1, mtrack_2, mtrack_3, mtrack_4, mtrack_5, mtrack_6, mtrack_7)
mako_tracks$species <- "I.oxyrinchus"

# blues
for(i in 1:length(blues)) {
  etuff <- read_etuff(blues[i])
  track <- get_track(etuff)
  track$ptt <- etuff$meta$ptt
  track <- track %>% separate(DateTime, into = c("Date", "Time"), sep = "([ ])")
  track <- track %>% mutate(
    kode = paste(Date, ptt, sep = "_")
  )
  assign(paste("btrack",i, sep = "_" ), track)
  rm(etuff, track)
  print(i)
}

blue_tracks <- rbind(btrack_1, btrack_2, btrack_3, btrack_4, btrack_5, btrack_6, btrack_7, btrack_8, btrack_9, btrack_10, btrack_11, btrack_12, btrack_13)
blue_tracks$species <- "P.glauca"

r <- c("/Volumes/A_DRIVE/bathy/northwest_atlantic.nc")
r <- raster(r)
r <- raster::aggregate(r, fact = 5)
r_df <- as.data.frame(r, xy = TRUE)
r_df <- r_df %>% filter(Elevation.relative.to.sea.level < 0)

hue_pal()(2)

ggplot(data = r_df) +
  geom_raster(aes(x = x, y = y, fill = Elevation.relative.to.sea.level), show.legend = FALSE) +
  scale_fill_gradient(low = "black", high = "grey") +
  coord_quickmap() +
  
  geom_path(data = btrack_1, aes(longitude, latitude), color = "#00BFC4") +
  geom_path(data = btrack_2, aes(longitude, latitude), color = "#00BFC4") +
  geom_path(data = btrack_3, aes(longitude, latitude), color = "#00BFC4") +
  geom_path(data = btrack_4, aes(longitude, latitude), color = "#00BFC4") +
  geom_path(data = btrack_5, aes(longitude, latitude), color = "#00BFC4") +
  geom_path(data = btrack_6, aes(longitude, latitude), color = "#00BFC4") +
  geom_path(data = btrack_7, aes(longitude, latitude), color = "#00BFC4") +
  geom_path(data = btrack_8, aes(longitude, latitude), color = "#00BFC4") +
  geom_path(data = btrack_9, aes(longitude, latitude), color = "#00BFC4") +
  geom_path(data = btrack_10, aes(longitude, latitude), color = "#00BFC4") +
  geom_path(data = btrack_11, aes(longitude, latitude), color = "#00BFC4") +
  geom_path(data = btrack_12, aes(longitude, latitude), color = "#00BFC4") +
  geom_path(data = btrack_13, aes(longitude, latitude), color = "#00BFC4") +
  
  geom_point(data = btrack_1[1,], aes(longitude, latitude), color = "green", shape = 1, size = 3) + 
  geom_point(data = btrack_2[1,], aes(longitude, latitude), color = "green", shape = 1, size = 3) + 
  geom_point(data = btrack_3[1,], aes(longitude, latitude), color = "green", shape = 1, size = 3) + 
  geom_point(data = btrack_4[1,], aes(longitude, latitude), color = "green", shape = 1, size = 3) + 
  geom_point(data = btrack_5[1,], aes(longitude, latitude), color = "green", shape = 1, size = 3) + 
  geom_point(data = btrack_6[1,], aes(longitude, latitude), color = "green", shape = 1, size = 3) +
  geom_point(data = btrack_7[1,], aes(longitude, latitude), color = "green", shape = 1, size = 3) + 
  geom_point(data = btrack_8[1,], aes(longitude, latitude), color = "green", shape = 1, size = 3) + 
  geom_point(data = btrack_9[1,], aes(longitude, latitude), color = "green", shape = 1, size = 3) + 
  geom_point(data = btrack_10[1,], aes(longitude, latitude), color = "green", shape = 1, size = 3) + 
  geom_point(data = btrack_11[1,], aes(longitude, latitude), color = "green", shape = 1, size = 3) + 
  geom_point(data = btrack_12[1,], aes(longitude, latitude), color = "green", shape = 1, size = 3) +
  geom_point(data = btrack_13[1,], aes(longitude, latitude), color = "green", shape = 1, size = 3) +
  
  geom_point(data = btrack_1[nrow(btrack_1),], aes(longitude, latitude), color = "red", shape = 4, size = 3) +
  geom_point(data = btrack_2[nrow(btrack_2),], aes(longitude, latitude), color = "red", shape = 4, size = 3) +
  geom_point(data = btrack_3[nrow(btrack_3),], aes(longitude, latitude), color = "red", shape = 4, size = 3) +
  geom_point(data = btrack_4[nrow(btrack_4),], aes(longitude, latitude), color = "red", shape = 4, size = 3) +
  geom_point(data = btrack_5[nrow(btrack_5),], aes(longitude, latitude), color = "red", shape = 4, size = 3) +
  geom_point(data = btrack_6[nrow(btrack_6),], aes(longitude, latitude), color = "red", shape = 4, size = 3) +
  geom_point(data = btrack_7[nrow(btrack_7),], aes(longitude, latitude), color = "red", shape = 4, size = 3) +
  geom_point(data = btrack_8[nrow(btrack_8),], aes(longitude, latitude), color = "red", shape = 4, size = 3) +
  geom_point(data = btrack_9[nrow(btrack_9),], aes(longitude, latitude), color = "red", shape = 4, size = 3) +
  geom_point(data = btrack_10[nrow(btrack_10),], aes(longitude, latitude), color = "red", shape = 4, size = 3) +
  geom_point(data = btrack_11[nrow(btrack_11),], aes(longitude, latitude), color = "red", shape = 4, size = 3) +
  geom_point(data = btrack_12[nrow(btrack_12),], aes(longitude, latitude), color = "red", shape = 4, size = 3) +
  geom_point(data = btrack_13[nrow(btrack_13),], aes(longitude, latitude), color = "red", shape = 4, size = 3) +
  
  geom_path(data = mtrack_1, aes(longitude, latitude), color = "#F8766D") +
  geom_path(data = mtrack_2, aes(longitude, latitude), color = "#F8766D") +
  geom_path(data = mtrack_3, aes(longitude, latitude), color = "#F8766D") +
  geom_path(data = mtrack_4, aes(longitude, latitude), color = "#F8766D") +
  geom_path(data = mtrack_5, aes(longitude, latitude), color = "#F8766D") +
  geom_path(data = mtrack_6, aes(longitude, latitude), color = "#F8766D") +
  geom_path(data = mtrack_7, aes(longitude, latitude), color = "#F8766D") +

  
  geom_point(data = mtrack_1[1,], aes(longitude, latitude), color = "green", shape = 1, size = 3) + 
  geom_point(data = mtrack_2[1,], aes(longitude, latitude), color = "green", shape = 1, size = 3) + 
  geom_point(data = mtrack_3[1,], aes(longitude, latitude), color = "green", shape = 1, size = 3) + 
  geom_point(data = mtrack_4[1,], aes(longitude, latitude), color = "green", shape = 1, size = 3) + 
  geom_point(data = mtrack_5[1,], aes(longitude, latitude), color = "green", shape = 1, size = 3) + 
  geom_point(data = mtrack_6[1,], aes(longitude, latitude), color = "green", shape = 1, size = 3) +
  geom_point(data = mtrack_7[1,], aes(longitude, latitude), color = "green", shape = 1, size = 3) +

  
  geom_point(data = mtrack_1[nrow(mtrack_1),], aes(longitude, latitude), color = "red", shape = 4, size = 3) +
  geom_point(data = mtrack_2[nrow(mtrack_2),], aes(longitude, latitude), color = "red", shape = 4, size = 3) +
  geom_point(data = mtrack_3[nrow(mtrack_3),], aes(longitude, latitude), color = "red", shape = 4, size = 3) +
  geom_point(data = mtrack_4[nrow(mtrack_4),], aes(longitude, latitude), color = "red", shape = 4, size = 3) +
  geom_point(data = mtrack_5[nrow(mtrack_5),], aes(longitude, latitude), color = "red", shape = 4, size = 3) +
  geom_point(data = mtrack_6[nrow(mtrack_6),], aes(longitude, latitude), color = "red", shape = 4, size = 3) +
  geom_point(data = mtrack_7[nrow(mtrack_7),], aes(longitude, latitude), color = "red", shape = 4, size = 3) +

  
  xlab("Longitude") +
  ylab("Latitude") +
  theme_minimal() +
  theme(element_blank())
```
``` {r cleanup}
rm(mtrack_1, mtrack_2, mtrack_3, mtrack_4, mtrack_5, mtrack_6, btrack_1, btrack_2, btrack_3, btrack_4, btrack_5, btrack_6, btrack_7, btrack_8, btrack_9, btrack_10, btrack_11, btrack_12, btrack_13)
```

# MAKING DATA SHOWCASE - SERIES TRACK OF BLUE SHARK DEPTH USE
``` {r SHOWCASE - 1 BLUE TRACK DAY}
# USING TRACKING DAY 2016-01-03 FROM BLUE SHARK 01

test <- bseries_7
test$duplicate <- test$DateTime_local
test <- test %>% separate(duplicate, into = c("Date_Local", "Local_Time"), sep = "([ ])")
test <- test %>% filter(Date == "2016-10-24")

test <- inner_join(test,pdtb) 


ggplot(data = test) + 
  geom_point(aes(x = as_hms(Time), y = depth, color = temperature)) +
  scale_color_viridis(option = "plasma") +
  scale_y_reverse() +
  scale_x_continuous(breaks = c(0, 43200, 86400), labels = c("00:00:00", "12:00:00", "23:00:00")) +
  ylab("Depth") +
  xlab("Time") +
  theme_classic()

# TAD SUMMARY FOR THIS SERIES DATA - THIS IS FOR PRESENTATION PURPOSES ONLY BECAUSE THE VALUES OF THE LAST BIN ARE EDITED
test <- omega_combo %>% filter(ptt.x == 141254 & Date.x == "2016-01-01")
formated <- data.frame(
  depth = c("0-10", "11-50", "51-100", "101-200", "201-300", "301-400", "401-500", "501-2000"),
  bin = c(rep(1:8, each = 1)),
  frequency = c(17.9,	13.7,	11.6,	16.8,	23.2,	3.2,	8.4,	0)
)


ggplot(data = formated, aes(x = bin, y = frequency)) +
  geom_col(fill = "grey22", color = "grey8", position = "dodge") +
  xlab("Depth (m)") +
  scale_x_reverse(breaks = c(1, 2, 3, 4, 5, 6, 7, 8), labels = c("0-10", "11-50", "51-100", "101-200", "201-300", "301-400", "401-500", "501-2000")) +
  ylab("Proportion of Time") +
  coord_flip() +
  theme_classic()
```

## MAKING HIGH-RESOLUTION CLUSTERING DATASET ##
```{r High res. cluster}
# adjust the time resolution of tag 206771 to match others 
toDelete <- c(seq(0,nrow(mseries_7), 2))
mseries_7.short <- mseries_7[-toDelete,]
rm(toDelete)

mako_series <- rbind(mseries_4, mseries_5, mseries_6, mseries_7.short)
combo_series <- rbind(mako_series, blue_series)

# filter out tracking days with less than 75% of their data
high_res <- combo_series %>% group_by(ptt) %>% count(Date)
high_res <- high_res %>% filter(n >= 345) # there are 675 (928) tracking days with more than 75% (60%) of their series data

high_res <- left_join(combo_series, high_res, by = c("Date", "ptt"))
high_res <- high_res %>% filter(!is.na(n))
# back-check
unique(high_res$kode) # vector of length 675
high_res %>% filter(species == "I.oxyrinchus") %>% count(kode) # 157 (194) tracking days
high_res %>% filter(species == "P.glauca") %>% count(kode) # 518 (734) tracking days

# count the net observations occur in each depth bin for day and night time
high_res <- high_res %>% group_by(ptt, Date, dn) 
daily.count <- high_res %>% count(Date)
high_res <- high_res %>% summarise(
  b1 = sum((depth < 10)),
  b2 = sum((depth >= 10 & depth <50)),
  b3 = sum((depth >= 50 & depth<100)),
  b4 = sum((depth >= 100 & depth<200)),
  b5 = sum((depth >= 200 & depth<300)),
  b6 = sum((depth >= 300 & depth<400)),
  b7 = sum((depth >= 400 & depth<500)),
  b8 = sum((depth >= 500)),
  sd = sd(depth),
  total = sum(b1,b2,b3,b4,b5,b6,b7,b8)
)

# check that sum of rows matches daily.count values
high_res <- ungroup(high_res)
test3 <- high_res %>% dplyr::select(ptt, Date, dn, total)

sum(daily.count$n != test3$total) # those look identical - moving on!
rm(test3)
# separate day and night time observations for each tracking day
daytime <- high_res %>% filter(dn == "d")
nightime <- high_res %>% filter(dn == "n")

# convert net counts into percentages within each depth bin
daytime <- daytime %>% transmute(
  ptt = ptt,
  Date = Date,
  d.b1 = (b1/total) * 100,
  d.b2 = (b2/total) * 100,
  d.b3 = (b3/total) * 100,
  d.b4 = (b4/total) * 100,
  d.b5 = (b5/total) * 100,
  d.b6 = (b6/total) * 100,
  d.b7 = (b7/total) * 100,
  d.b8 = (b8/total) * 100,
  d.sd = sd
)

nightime <- nightime %>% transmute(
  ptt = ptt,
  Date = Date,
  n.b1 = (b1/total) * 100,
  n.b2 = (b2/total) * 100,
  n.b3 = (b3/total) * 100,
  n.b4 = (b4/total) * 100,
  n.b5 = (b5/total) * 100,
  n.b6 = (b6/total) * 100,
  n.b7 = (b7/total) * 100,
  n.b8 = (b8/total) * 100,
  n.sd = sd
)

high_res <- inner_join(daytime, nightime, by = c("ptt", "Date"))
rm(daytime, nightime)

# now create a species stamp to apply to this object
species_stmp <- data.frame(
  ptt = c(141257, 163098, 163096, 141254, 141256, 141259, 106754, 133016, 133017, 133018, 133021, 141247, 141255, 163097, 141258, 154096, 78680,  78682,  78683, 206771),
  species = c("I.oxyrinchus", "I.oxyrinchus", "I.oxyrinchus", "P.glauca", "P.glauca", "P.glauca", "P.glauca", "P.glauca", "P.glauca", "P.glauca", "P.glauca", "P.glauca", "P.glauca", "P.glauca", "P.glauca", "P.glauca", "I.oxyrinchus", "I.oxyrinchus", "I.oxyrinchus", "I.oxyrinchus")
)
high_res <- left_join(high_res, species_stmp, by = c("ptt"))

# add kode to high_res
high_res <- high_res %>% mutate(
  kode = paste(Date, ptt, sep = "_")
)

# add depth over location of each tracking day to high-resolution
high_res <- inner_join(high_res, bathy_stamp, by = c("kode"))
```

## CLUSTER ANALYSIS 2.0: HIGH RESOLUTION TAD SUMMARIES ##
This cluster anlysis is performed using the day/night summaries of time-at-depth with the further addition of depth standard deviation as a proxy for total vertical movement. These data are compiled from time-series depth information in chunk #15: 

# Step 1: run the clusters with each species - can include species as a cluster component or not
``` {r Heirarchical clustering}
# remove tracking days over water more shallow than 1000 meters
Ctad2 <- high_res %>% filter(depth <= -1000)
# Create a dataframe of only cluster variables
Ctad2 <- Ctad2 %>% dplyr::select(-c("ptt", "Date", "species", "kode", "depth"))

## Heirarchical Clustering
# create a distance matrix
combo_dist2 <- dist(Ctad2, method = "manhattan")
# perform cluster assignment
Clust2 <- hclust(combo_dist2, method = "average")
```

# Step 2: Assessing the statistical validity of the clusters
``` {r NbClust tests}
# CH INDEX:
NbClust(data = Ctad2, diss = NULL, distance = "manhattan", min.nc = 2,
        max.nc = 20, method = "average", index = "ch")
# Raw: 10

# DUDA TEST: Smallest number of clusters which produces a critical value greater than the Duda score
NbClust(data = Ctad2, diss = NULL, distance = "manhattan", min.nc = 2,
        max.nc = 20, method = "average", index = "duda")
# Raw: 2 clusters 

# C INDEX
NbClust(data = Ctad2, diss = NULL, distance = "manhattan", min.nc = 2,
        max.nc = 20, method = "average", index = "cindex")
# Raw: 7 clusters 

# BEALE INDEX
NbClust(data = Ctad2, diss = NULL, distance = "manhattan", min.nc = 2,
        max.nc = 20, method = "average", index = "beale")
# Raw: 2 clusters 
```

Step 3: partition the data into its rightful cluster
```{r fine-scale cluster assignment}
combo_sub_grp2 <- cutree(Clust2, k = 11)
table(combo_sub_grp2)

Ctad2$cluster <- combo_sub_grp2

high_res$cluster <- combo_sub_grp2
clust_stamp2 <- high_res %>% dplyr::select(kode, species, cluster)
```

## CLUSTER COMPARISONS
# TAD profiles
Create basic histograms of cluster tad profiles 
``` {r TAD profiles}
# create histograms to visualize the clusters
daytime.combo_summ <- Ctad2 %>%
  group_by(cluster) %>%
  summarize(
    d.one = mean(d.b1),
    d.two = mean(d.b2),
    d.three = mean(d.b3),
    d.four = mean(d.b4),
    d.five = mean(d.b5),
    d.six = mean(d.b6),
    d.seven = mean(d.b7),
    d.eight = mean(d.b8),
  )
daytime.combo_summ <- dplyr::select(combo_summ, -c(cluster))

format_tad_combo <- data.frame(
  depth = c(rep(bins_max, times = 11)),
  bin = c(rep(1:8, times = 11)),
  frequency = c(as.numeric(daytime.combo_summ[1,]),
                as.numeric(daytime.combo_summ[2,]),
                as.numeric(daytime.combo_summ[3,]),
                as.numeric(daytime.combo_summ[4,]),
                as.numeric(daytime.combo_summ[5,]),
                as.numeric(daytime.combo_summ[6,]),
                as.numeric(daytime.combo_summ[7,]),
                as.numeric(daytime.combo_summ[8,]),
                as.numeric(daytime.combo_summ[9,]),
                as.numeric(daytime.combo_summ[10,])
                as.numeric(daytime.combo_summ[11,])),
  cluster = c(rep(1:11, each = 8))
)


# CLUSTER 1

format_tad_combo %>%
  filter(cluster == 1) %>%
  ggplot(aes(x = bin, y = frequency)) +
  geom_col(fill = "grey22", color = "grey8", position = "dodge") +
  ggtitle("Combo Shark Cluster 1") +
  xlab("Depth") +
  ylab("Proportion of Occupancy") +
  scale_x_reverse(breaks = c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.5, 7.5, 8.5), labels = c("0", "10", "50", "100", "200", "300", "400", "500", "2000")) +
  coord_flip() +
  theme_classic()

cluster_species_freq %>% filter(cluster == 1) %>% 
  ggplot() +
  geom_col(aes(x = by, y = normalized, color = by, fill = by)) +
  ggtitle("Combo Shark Cluster 1 Observations by Species") +
  coord_flip() +
  theme_classic()

# CLUSTER 2

format_tad_combo %>%
  filter(cluster == 2) %>%
  ggplot(aes(x = bin, y = frequency)) +
  geom_col(fill = "grey22", color = "grey8", position = "dodge") +
  ggtitle("Combo Shark Cluster 2") +
  xlab("Depth") +
  ylab("Proportion of Occupancy") +
  scale_x_reverse(breaks = c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.5, 7.5, 8.5), labels = c("0", "10", "50", "100", "200", "300", "400", "500", "2000")) +
  coord_flip() +
  theme_classic()

cluster_species_freq %>% filter(cluster == 2) %>% 
  ggplot() +
  geom_col(aes(x = by, y = normalized, color = by, fill = by)) +
  ggtitle("Combo Shark Cluster 2 Observations by Species") +
  coord_flip() +
  theme_classic()

# CLUSTER 3

format_tad_combo %>%
  filter(cluster == 3) %>%
  ggplot(aes(x = bin, y = frequency)) +
  geom_col(fill = "grey22", color = "grey8", position = "dodge") +
  ggtitle("Combo Shark Cluster 3") +
  xlab("Depth") +
  ylab("Proportion of Occupancy")
  scale_x_reverse(breaks = c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.5, 7.5, 8.5), labels = c("0", "10", "50", "100", "200", "300", "400", "500", "2000")) +
  coord_flip() +
  theme_classic()

cluster_species_freq %>% filter(cluster == 3) %>% 
  ggplot() +
  geom_col(aes(x = by, y = normalized, color = by, fill = by)) +
  ggtitle("Combo Shark Cluster 3 Observations by Species") +
  coord_flip() +
  theme_classic()

# CLUSTER 4

format_tad_combo %>%
  filter(cluster == 4) %>%
  ggplot(aes(x = bin, y = frequency)) +
  geom_col(fill = "grey22", color = "grey8", position = "dodge") +
  ggtitle("Combo Shark Cluster 4") +
  xlab("Depth") +
  ylab("Proportion of Occupancy")
  scale_x_reverse(breaks = c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.5, 7.5, 8.5), labels = c("0", "10", "50", "100", "200", "300", "400", "500", "2000")) +
  coord_flip() +
  theme_classic()

cluster_species_freq %>% filter(cluster == 4) %>% 
  ggplot() +
  geom_col(aes(x = by, y = normalized, color = by, fill = by)) +
  ggtitle("Combo Shark Cluster 4 Observations by Species") +
  coord_flip() +
  theme_classic()

# CLUSTER 5

format_tad_combo %>%
  filter(cluster == 5) %>%
  ggplot(aes(x = bin, y = frequency)) +
  geom_col(fill = "grey22", color = "grey8", position = "dodge") +
  ggtitle("Combo Shark Cluster 5") +
  xlab("Depth") +
  ylab("Proportion of Occupancy")
  scale_x_reverse(breaks = c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.5, 7.5, 8.5), labels = c("0", "10", "50", "100", "200", "300", "400", "500", "2000")) +
  coord_flip() +
  theme_classic()

cluster_species_freq %>% filter(cluster == 5) %>% 
  ggplot() +
  geom_col(aes(x = by, y = normalized, color = by, fill = by)) +
  ggtitle("Combo Shark Cluster 5 Observations by Species") +
  coord_flip() +
  theme_classic()

# CLUSTER 6

format_tad_combo %>%
  filter(cluster == 6) %>%
  ggplot(aes(x = bin, y = frequency)) +
  geom_col(fill = "grey22", color = "grey8", position = "dodge") +
  ggtitle("Combo Shark Cluster 6") +
  xlab("Depth") +
  ylab("Proportion of Occupancy")
  scale_x_reverse(breaks = c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.5, 7.5, 8.5), labels = c("0", "10", "50", "100", "200", "300", "400", "500", "2000")) +
  coord_flip() +
  theme_classic()

cluster_species_freq %>% filter(cluster == 6) %>% 
  ggplot() +
  geom_col(aes(x = by, y = normalized, color = by, fill = by)) +
  ggtitle("Combo Shark Cluster 6 Observations by Species") +
  coord_flip() +
  theme_classic()

# CLUSTER 7

format_tad_combo %>%
  filter(cluster == 7) %>%
  ggplot(aes(x = bin, y = frequency)) +
  geom_col(fill = "grey22", color = "grey8", position = "dodge") +
  ggtitle("Combo Shark Cluster 7") +
  xlab("Depth") +
  ylab("Proportion of Occupancy")
  scale_x_reverse(breaks = c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.5, 7.5, 8.5), labels = c("0", "10", "50", "100", "200", "300", "400", "500", "2000")) +
  coord_flip() +
  theme_classic()

cluster_species_freq %>% filter(cluster == 7) %>% 
  ggplot() +
  geom_col(aes(x = by, y = normalized, color = by, fill = by)) +
  ggtitle("Combo Shark Cluster 7 Observations by Species") +
  coord_flip() +
  theme_classic()

# CLUSTER 8

format_tad_combo %>%
  filter(cluster == 8) %>%
  ggplot(aes(x = bin, y = frequency)) +
  geom_col(fill = "grey22", color = "grey8", position = "dodge") +
  ggtitle("Combo Shark Cluster 8") +
  xlab("Depth") +
  ylab("Proportion of Occupancy")
  scale_x_reverse(breaks = c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.5, 7.5, 8.5), labels = c("0", "10", "50", "100", "200", "300", "400", "500", "2000")) +
  coord_flip() +
  theme_classic()

cluster_species_freq %>% filter(cluster == 8) %>% 
  ggplot() +
  geom_col(aes(x = by, y = normalized, color = by, fill = by)) +
  ggtitle("Combo Shark Cluster 8 Observations by Species") +
  coord_flip() +
  theme_classic()

# CLUSTER 9

format_tad_combo %>%
  filter(cluster == 9) %>%
  ggplot(aes(x = bin, y = frequency)) +
  geom_col(fill = "grey22", color = "grey8", position = "dodge") +
  ggtitle("Combo Shark Cluster 9") +
  xlab("Depth") +
  ylab("Proportion of Occupancy")
  scale_x_reverse(breaks = c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.5, 7.5, 8.5), labels = c("0", "10", "50", "100", "200", "300", "400", "500", "2000")) +
  coord_flip() +
  theme_classic()

cluster_species_freq %>% filter(cluster == 9) %>% 
  ggplot() +
  geom_col(aes(x = by, y = normalized, color = by, fill = by)) +
  ggtitle("Combo Shark Cluster 9 Observations by Species") +
  coord_flip() +
  theme_classic()
```

# Omega Combo
This chunk creates the omega_combo data frame with environmental data for each tracking day. 
```{r omega combo}
tad_combo <- tad_combo %>% mutate(
  kode = paste(Date, ptt, sep = "_")
)

omega_mako <- read_excel("~/Desktop/blues-makos/data/omega_mako.xlsx")
omega_blue <- read_excel("~/Desktop/blues-makos/data/omega_blue.xlsx")

omega_mako <- omega_mako %>% dplyr::select(c(kode, Date.y, Time, localHour.y, x, y, sst, sst_sd, ssh, ssh_sd, eke, ild.5, n2, lunar, bathy, chloro))
omega_blue <- omega_blue %>% dplyr::select(c(kode, Date.y, Time, localHour.y, x, y, sst, sst_sd, ssh, ssh_sd, eke, ild.5, n2, lunar, bathy, chloro))

omega_combo <- rbind(omega_mako, omega_blue)

omega_combo <- omega_combo %>% separate(kode, into = c("Date", "ptt"), sep = "([ ])")
omega_combo <- omega_combo %>% mutate(
  kode = paste(Date, ptt, sep = "_")
)
omega_combo <- omega_combo %>% dplyr::select(-c(Date, ptt))

omega_combo <- inner_join(omega_combo, tad_combo, by = c("kode"))
omega_combo$cluster <- as.factor(omega_combo$cluster)
omega_combo$ptt <- as.factor(omega_combo$ptt)

omega_combo <- omega_combo %>% separate(Date.y, into = c("Year", "Month", "Day"), sep = "([-])")

```

# Map of Cluster Locations
Across the North Atlantic study area, where did each cluster occur along the track of each individual shark?
```{r cluster maps}
combo_track <- rbind(mako_tracks, blue_tracks)
combo_track <- inner_join(combo_track, clust_stamp2, by = c("kode"))
combo_track$cluster <- factor(combo_track$cluster)
combo_track$species <- factor(combo_track$species.x)

mako_tracks <- inner_join(mako_tracks, clust_stamp)
blue_tracks <- inner_join(blue_tracks, clust_stamp)

# bathymetry raster layer
b <- c("/Volumes/A_DRIVE/bathy/bathy.nc")
b <- raster(r)
b <- raster::aggregate(r, fact = 5)
b_df <- as.data.frame(r, xy = TRUE)
b_df <- r_df %>% filter(Elevation.relative.to.sea.level < 0)

# bathymetry
ggplot() +
  geom_raster(data = filter(r_df, Elevation.relative.to.sea.level < 0), aes(x = x, y = y, fill = Elevation.relative.to.sea.level)) +
  geom_sf(data = world, color = "black") +
  coord_sf(xlim = c(-80, -10), ylim = c(7,47), expand = FALSE) +
  geom_point(data = combo_track, aes(longitude, latitude, color = species)) +
  facet_wrap(~species) +
  xlab("Longitude") +
  ylab("Latitude") +
  theme_minimal() +
  theme(legend.position = "none")

# cluster 1
ggplot() +
  geom_raster(data = filter(r_df, Elevation.relative.to.sea.level < 0), aes(x = x, y = y, fill = Elevation.relative.to.sea.level)) +
  scale_fill_gradient(low = "black", high = "grey") +
  geom_sf(data = world, color = "black") +
  coord_sf(xlim = c(-80, -10), ylim = c(7,47), expand = FALSE) +
  geom_point(data = filter(combo_track, cluster == 1), aes(longitude, latitude, color = species.x)) +
  xlab("Longitude") +
  ylab("Latitude") +
  ggtitle("Cluster 1") +
  theme_minimal() +
  theme(legend.position = "none")

# cluster 2
ggplot() +
  geom_raster(data = filter(r_df, Elevation.relative.to.sea.level < 0), aes(x = x, y = y, fill = Elevation.relative.to.sea.level)) +
  scale_fill_gradient(low = "black", high = "grey") +
  geom_sf(data = world, color = "black") +
  coord_sf(xlim = c(-80, -10), ylim = c(7,47), expand = FALSE) +
  geom_point(data = filter(combo_track, cluster == 2), aes(longitude, latitude, color = species.x)) +
  xlab("Longitude") +
  ylab("Latitude") +
  ggtitle("Cluster 2") +
  theme_minimal() +
  theme(legend.position = "none")

# cluster 3
ggplot() +
  geom_raster(data = filter(r_df, Elevation.relative.to.sea.level < 0), aes(x = x, y = y, fill = Elevation.relative.to.sea.level)) +
  scale_fill_gradient(low = "black", high = "grey") +
  geom_sf(data = world, color = "black") +
  coord_sf(xlim = c(-80, -10), ylim = c(7,47), expand = FALSE) +
  geom_point(data = filter(combo_track, cluster == 3), aes(longitude, latitude, color = species.x)) +
  xlab("Longitude") +
  ylab("Latitude") +
  ggtitle("Cluster 3") +
  theme_minimal() +
  theme(legend.position = "none")

# cluster 4
ggplot() +
  geom_raster(data = filter(r_df, Elevation.relative.to.sea.level < 0), aes(x = x, y = y, fill = Elevation.relative.to.sea.level)) +
  scale_fill_gradient(low = "black", high = "grey") +
  geom_sf(data = world, color = "black") +
  coord_sf(xlim = c(-80, -10), ylim = c(7,47), expand = FALSE) +
  geom_point(data = filter(combo_track, cluster == 4), aes(longitude, latitude, color = species.x)) +
  xlab("Longitude") +
  ylab("Latitude") +
  ggtitle("Cluster 4") +
  theme_minimal() +
  theme(legend.position = "none")

# cluster 5
ggplot() +
  geom_raster(data = filter(r_df, Elevation.relative.to.sea.level < 0), aes(x = x, y = y, fill = Elevation.relative.to.sea.level)) +
  scale_fill_gradient(low = "black", high = "grey") +
  geom_sf(data = world, color = "black") +
  coord_sf(xlim = c(-80, -10), ylim = c(7,47), expand = FALSE) +
  geom_point(data = filter(combo_track, cluster == 5), aes(longitude, latitude, color = species.x)) +
  xlab("Longitude") +
  ylab("Latitude") +
  ggtitle("Cluster 5") +
  theme_minimal() +
  theme(legend.position = "none")

```

# Dendrograms
Create a dendrogram showing the separation of the different heirarchical clusters
Note: PRUNED DENDROGRAM CODE DOES NOT SEEM TO BE FUNCTIONING PROPERLY IN ITS CURRENT STATE
```{r dendrogram}
plot(Clust2)
# trying to figure out what parts of the dendrogram corespond to which clusters
cluster_dendro <- as.dendrogram(Clust2)
colors_to_use <- as.numeric(Ctad2$cluster)
colors_to_use <- colors_to_use[order.dendrogram(cluster_dendro)]
labels_colors(cluster_dendro) <- colors_to_use
labels_colors(cluster_dendro)
plot(cluster_dendro)

# make a list of ggplot colors
hue_pal()(10)
show_col(hue_pal()(10))
ggplot_colors <- data.frame(
  numb = c(1,2,3,4,5,6,7,8,9,10),
  code = c("#F8766D", "#D89000", "#A3A500", "#39B600", "#00BF7D", "#00BFC4", "#00B0F6", "#9590FF", "#E76BF3", "#FF62BC")
)

# finally, make a pretty colored dendrogram
color_dendro <- color_branches(cluster_dendro, clusters = colors_to_use, col = c("#00B0F6", "#00BFC4", "#FF62BC", "#E76BF3", "#F8766D", "#D89000", "#9590FF", "#39B600", "#A3A500", "#00BF7D"))

color_dendro <- color_branches(cluster_dendro, clusters = colors_to_use, col = c("black", "black", "black", "black", "#F8766D", "#F8766D", "#00BFC4", "#00BFC4", "#00BFC4", "#00BFC4")) # general patterns
plot(color_dendro)
```

# Heatmaps ## NEEDS UPDATING ##
Create heatmaps of vertical distribution within each cluster. Here, each column represents a destinct tracking day with different depth bins extending down into the water column along the y-axis. The color of each box represents the proportion of time that the animal spent in that depth bin on that particular day. 
``` {r cluster heatmaps}
# makos
combo_heatmap <- tad_combo %>% dplyr::select(-c(Date, ptt, owner, kode, species, cluster))
combo_heatmap$Row <- (1:nrow(combo_heatmap))

pived_stamp <- data.frame(
  species = rep(tad_combo$species, each = 8),
  cluster = rep(tad_combo$cluster, each = 8)
)

pived <- combo_heatmap %>% 
  pivot_longer( -Row, names_to = "bin", values_to = "count")

pived <- cbind(pived, pived_stamp)

pived$bin <- factor(pived$bin, levels = c("bin1", "bin2", "bin3", "bin4", "bin5", "bin6", "bin7", "bin8"), ordered = TRUE)
pived$bin <- as.numeric(pived$bin)
pived$Row <- factor(pived$Row, levels = c(unique(combo_heatmap$Row)), ordered = TRUE)

# repeat for all 9 clusters and both species
# MAKOS
# C1
ggplot(data = filter(pived, species == "I. oxyrinchus")) +
  geom_tile(data = filter(pived, species == "I. oxyrinchus" & cluster ==1), aes(x = Row, y = bin, fill = count), linejoin = "round") +
  scale_fill_viridis(limits = c(0,100), discrete = FALSE) +
  theme_dendro() +
  scale_y_reverse(breaks = c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.5, 7.5, 8.5), labels = c("0", "10", "50", "100", "200", "300", "400", "500", "2000")) +
  theme(aspect.ratio = 20/10)
# C2
ggplot(data = filter(pived, species == "I. oxyrinchus")) +
  geom_tile(data = filter(pived, species == "I. oxyrinchus" & cluster ==2), aes(x = Row, y = bin, fill = count), linejoin = "round") +
  scale_fill_viridis(limits = c(0,100), discrete = FALSE) +
  theme_dendro() +
  scale_y_reverse(breaks = c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.5, 7.5, 8.5), labels = c("0", "10", "50", "100", "200", "300", "400", "500", "2000")) +
  theme(aspect.ratio = 20/10)
# C3
ggplot(data = filter(pived, species == "I. oxyrinchus")) +
  geom_tile(data = filter(pived, species == "I. oxyrinchus" & cluster ==3), aes(x = Row, y = bin, fill = count), linejoin = "round") +
  scale_fill_viridis(limits = c(0,100), discrete = FALSE) +
  theme_dendro() +
  scale_y_reverse(breaks = c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.5, 7.5, 8.5), labels = c("0", "10", "50", "100", "200", "300", "400", "500", "2000")) +
  theme(aspect.ratio = 20/10)
 # C4
ggplot(data = filter(pived, species == "I. oxyrinchus")) +
  geom_tile(data = filter(pived, species == "I. oxyrinchus" & cluster ==4), aes(x = Row, y = bin, fill = count), linejoin = "round") +
  scale_fill_viridis(limits = c(0,100), discrete = FALSE) +
  theme_dendro() +
  scale_y_reverse(breaks = c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.5, 7.5, 8.5), labels = c("0", "10", "50", "100", "200", "300", "400", "500", "2000")) +
  theme(aspect.ratio = 20/10)
# C5
ggplot(data = filter(pived, species == "I. oxyrinchus")) +
  geom_tile(data = filter(pived, species == "I. oxyrinchus" & cluster ==5), aes(x = Row, y = bin, fill = count), linejoin = "round") +
  scale_fill_viridis(limits = c(0,100), discrete = FALSE) +
  theme_dendro() +
  scale_y_reverse(breaks = c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.5, 7.5, 8.5), labels = c("0", "10", "50", "100", "200", "300", "400", "500", "2000")) +
  theme(aspect.ratio = 20/10)
# C6
ggplot(data = filter(pived, species == "I. oxyrinchus")) +
  geom_tile(data = filter(pived, species == "I. oxyrinchus" & cluster ==6), aes(x = Row, y = bin, fill = count), linejoin = "round") +
  scale_fill_viridis(limits = c(0,100), discrete = FALSE) +
  theme_dendro() +
  scale_y_reverse(breaks = c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.5, 7.5, 8.5), labels = c("0", "10", "50", "100", "200", "300", "400", "500", "2000")) +
  theme(aspect.ratio = 20/10)
# C7
ggplot(data = filter(pived, species == "I. oxyrinchus")) +
  geom_tile(data = filter(pived, species == "I. oxyrinchus" & cluster ==7), aes(x = Row, y = bin, fill = count), linejoin = "round") +
  scale_fill_viridis(limits = c(0,100), discrete = FALSE) +
  theme_dendro() +
  scale_y_reverse(breaks = c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.5, 7.5, 8.5), labels = c("0", "10", "50", "100", "200", "300", "400", "500", "2000")) +
  theme(aspect.ratio = 20/10)
# C8
ggplot(data = filter(pived, species == "I. oxyrinchus")) +
  geom_tile(data = filter(pived, species == "I. oxyrinchus" & cluster ==8), aes(x = Row, y = bin, fill = count), linejoin = "round") +
  scale_fill_viridis(limits = c(0,100), discrete = FALSE) +
  theme_dendro() +
  scale_y_reverse(breaks = c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.5, 7.5, 8.5), labels = c("0", "10", "50", "100", "200", "300", "400", "500", "2000")) +
  theme(aspect.ratio = 20/10)
# C9
ggplot(data = filter(pived, species == "I. oxyrinchus")) +
  geom_tile(data = filter(pived, species == "I. oxyrinchus" & cluster ==9), aes(x = Row, y = bin, fill = count), linejoin = "round") +
  scale_fill_viridis(limits = c(0,100), discrete = FALSE) +
  theme_dendro() +
  scale_y_reverse(breaks = c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.5, 7.5, 8.5), labels = c("0", "10", "50", "100", "200", "300", "400", "500", "2000")) +
  theme(aspect.ratio = 20/10)

# BLUES
# C1
ggplot(data = filter(pived, species == "P. glauca")) +
  geom_tile(data = filter(pived, species == "P. glauca" & cluster ==1), aes(x = Row, y = bin, fill = count), linejoin = "round") +
  scale_fill_viridis(limits = c(0,100), discrete = FALSE) +
  theme_dendro() +
  scale_y_reverse(breaks = c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.5, 7.5, 8.5), labels = c("0", "10", "50", "100", "200", "300", "400", "500", "2000")) +
  theme(aspect.ratio = 20/10)
# C2
ggplot(data = filter(pived, species == "P. glauca")) +
  geom_tile(data = filter(pived, species == "P. glauca" & cluster ==2), aes(x = Row, y = bin, fill = count), linejoin = "round") +
  scale_fill_viridis(limits = c(0,100), discrete = FALSE) +
  theme_dendro() +
  scale_y_reverse(breaks = c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.5, 7.5, 8.5), labels = c("0", "10", "50", "100", "200", "300", "400", "500", "2000")) +
  theme(aspect.ratio = 20/10)
# C3
ggplot(data = filter(pived, species == "P. glauca")) +
  geom_tile(data = filter(pived, species == "P. glauca" & cluster ==3), aes(x = Row, y = bin, fill = count), linejoin = "round") +
  scale_fill_viridis(limits = c(0,100), discrete = FALSE) +
  theme_dendro() +
  scale_y_reverse(breaks = c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.5, 7.5, 8.5), labels = c("0", "10", "50", "100", "200", "300", "400", "500", "2000")) +
  theme(aspect.ratio = 20/10)
# C4
ggplot(data = filter(pived, species == "P. glauca")) +
  geom_tile(data = filter(pived, species == "P. glauca" & cluster ==4), aes(x = Row, y = bin, fill = count), linejoin = "round") +
  scale_fill_viridis(limits = c(0,100), discrete = FALSE) +
  theme_dendro() +
  scale_y_reverse(breaks = c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.5, 7.5, 8.5), labels = c("0", "10", "50", "100", "200", "300", "400", "500", "2000")) +
  theme(aspect.ratio = 20/10)
# C5
ggplot(data = filter(pived, species == "P. glauca")) +
  geom_tile(data = filter(pived, species == "P. glauca" & cluster ==5), aes(x = Row, y = bin, fill = count), linejoin = "round") +
  scale_fill_viridis(limits = c(0,100), discrete = FALSE) +
  theme_dendro() +
  scale_y_reverse(breaks = c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.5, 7.5, 8.5), labels = c("0", "10", "50", "100", "200", "300", "400", "500", "2000")) +
  theme(aspect.ratio = 20/10)
# C6
ggplot(data = filter(pived, species == "P. glauca")) +
  geom_tile(data = filter(pived, species == "P. glauca" & cluster ==6), aes(x = Row, y = bin, fill = count), linejoin = "round") +
  scale_fill_viridis(limits = c(0,100), discrete = FALSE) +
  theme_dendro() +
  scale_y_reverse(breaks = c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.5, 7.5, 8.5), labels = c("0", "10", "50", "100", "200", "300", "400", "500", "2000")) +
  theme(aspect.ratio = 20/10)
# C7
ggplot(data = filter(pived, species == "P. glauca")) +
  geom_tile(data = filter(pived, species == "P. glauca" & cluster ==7), aes(x = Row, y = bin, fill = count), linejoin = "round") +
  scale_fill_viridis(limits = c(0,100), discrete = FALSE) +
  theme_dendro() +
  scale_y_reverse(breaks = c(0.5, 1.5, 2.5, 3.5, 4.5, 5.5, 6.5, 7.5, 8.5), labels = c("0", "10", "50", "100", "200", "300", "400", "500", "2000")) +
  theme(aspect.ratio = 20/10)
```

# Proportion time at depth accross daylight hours heatmaps
This might work right out of the box with the series data_frame I made earlier...
Note: MAKE SURE TO USE LOCAL TIME

``` {r TAD daylight}
combo_series <- rbind(mako_series, blue_series)
combo_series$Local_Time <- as_hms(combo_series$Local_Time)
cluster_series <- inner_join(combo_series, clust_stamp2, by = "kode")
unique(combo_series$kode)

# daily cluster kernal density heatmaps
ggplot(data = filter(cluster_series, cluster == 1), aes(x = Local_Time, y = depth) ) +
  stat_density_2d(aes(fill = ((..ndensity..))), geom = "raster", contour = FALSE) +
  facet_wrap(~cluster) +
  scale_fill_viridis(option = "magma") + 
  labs(fill = "Density") +
  scale_y_reverse(breaks = c(0, 500, 1000, 1500), labels = c("0", "500", "1000", "1500")) +
  scale_x_continuous(breaks = c(0, 43200, 86400), labels = c("00:00:00", "12:00:00", "23:00:00")) +
  xlab("Time of Day") +
  ylab("Depth(m)") +
  theme_minimal()

# daily cluster point maps
ggplot(data = filter(cluster_series, cluster == 5)) +
  geom_point(aes(x = Local_Time, y = depth, color = species.x), alpha = 0.01) +
  facet_grid(cluster ~ species.x) + 
  scale_y_reverse(limits = c(1000, 0), breaks = c(1000, 500, 0)) +
  theme_classic() +
  theme(legend.position = "bottom")
```

# Series Plots - Mass Production
Generate plots of series data for all tracking days w/ more than 75% data coverage
``` {r series assemly line}
# update line 1481 and line 1505 with appropriate cluster number
c1 <- cluster_series %>% filter(cluster == 5)
series_dates <- unique(c1$Date) 

for(i in 1:length(series_dates)) {
  day <- c1 %>% filter(Date == series_dates[i])
  sharks <- day %>% count(ptt)
  sharks <- sharks %>% filter(n >= 260)
  ptt_n <- c(sharks$ptt)
  full_series <- data.frame()
  for(i in 1:length(ptt_n)) {
    tryCatch({
    x <- day %>% filter(ptt == ptt_n[i])
    full_series <- rbind(full_series, x)
    rm(x)
  }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
  }
 if (nrow(full_series) != 0) {
 ggplot() + 
    geom_line(data = filter(full_series, species.x == "P.glauca"), aes(x = DateTime, y = depth), color = "#00BFC4") +
    geom_line(data = filter(full_series, species.x == "I.oxyrinchus"), aes(x = DateTime, y = depth), color = "#F8766D") +
    scale_y_reverse() +
    facet_grid(ptt~.) +
    theme_classic()
  
  ggsave(paste("~/Desktop/blues-makos/images/cluster_5/", full_series[1,1], ".png", sep = ""))
 
  rm(day, sharks, ptt_n, full_series)
 }
  else {
    rm(day, sharks, ptt_n, full_series)
  }
}

```

## EXTRACTING ENVIRONMENTAL DATA ##
# Cam code
Here's a modified example of some extraction code which Cam wrote for me a while back:
```{r lucky mako 7}
# adding the 7th mako to omega_combo
# separate tad and track data

m7 <- tad_mako %>% filter(ptt == 206771)
m7 <- m7 %>% mutate(
  kode = paste(Date, ptt, sep = "_")
)

## get the track and trim to unique DAYS (not unique date-times as that just introduces duplication into the loop below)
etuff <- read_etuff(makos[7])
track7 <- get_track(etuff)
track7$dup <- track7$DateTime
track7 <- track7 %>% separate(dup, into = c("Date", "Time"), sep = "([ ])")
track7$ptt <- etuff$meta$ptt
track7 <- track7 %>% mutate(
  kode = paste(Date, ptt, sep = "_")
)
track7 <- track7 %>% dplyr::select(c("kode", "latitude", "longitude", "Date", "DateTime"))
track7$day <- as.Date(track7$DateTime, tz='UTC')
track7 <- track7[which(!duplicated(track7$day)),]
# now stick them together
m7 <- left_join(m7, track7)
```

``` {r HYCOM}
# create a stamp with coordinates for each tracking day
position.stmp <- combo_track %>% dplyr::select(latitude, longitude, kode)
position.stmp <- position.stmp[which(!duplicated(position.stmp$kode)),] # remove duplicates in days w. more than one location
# apply position data to high_res data frame
high_res <- inner_join(high_res, position.stmp, by = c("kode"))
# extract environmental files from HYCOM for each unique date
## create an index of new columns where you want the environmental data to go. I happen to know this outputs 15 cols so I cheated...
col_idx <- c((ncol(high_res) + 1):(ncol(high_res) + 16))

t1 <- Sys.time()
for (tt in 1:nrow(high_res)){
  data <- unlist(facet_hycom(xpos = high_res$longitude[tt],
                                        ypos = high_res$latitude[tt],
                                        tpos = as.Date(high_res$Date[tt]), ## needs to be of class 'Date'
                                        xlen = 0.25, ## these "errors" need to be approx this size to allow the calculations to run successfully
                                        ylen = 0.25, 
                                        varName = c('water_temp', 'water_u', 'water_v','surf_el','salinity')))
  data <- as.data.frame(t(data))
  high_res[tt,col_idx] <- data
  rm(data)
}

t2 <- Sys.time()
t2 - t1
```

```{r lucky mako 7: BATHYMETRY}
# extracting Bathymetry data
nc_data <- nc_open("/Volumes/A_DRIVE/bathy/northwest_atlantic.nc")

lon <- ncvar_get(nc_data, "lon")
lat <- ncvar_get(nc_data, "lat")

depth.array <- ncvar_get(nc_data, "elevation")
# find out what value is input for misssing data
fillvalue <- ncatt_get(nc_data, "elevation", "_FillValue")
# close the NETCDF file
nc_close(nc_data) 
# get model points for extraction from LiveOcean #
lon_psat <- m7$longitude
lat_psat <- m7$latitude

extraction_points <- data.frame (
  longitude = c(lon_psat),
  latitude = c(lat_psat)
)
# transform the array into a raster layer
r_brick <- raster(depth.array, xmn=min(lat), xmx=max(lat), ymn=min(lon), ymx=max(lon), crs=CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs+ towgs84=0,0,0"))
# check orrientation
plot(r_brick)
# correct the orrientation of the raster brick
r_brick <- flip(t(r_brick), direction='y')

# extracting points from the brick
bathy <- data.frame()

for(i in 1:nrow(extraction_points)) {
  point_series <- raster::extract(r_brick, extraction_points[i,], method='simple')
  data <- data.frame(
    point = c(i),
    lat = extraction_points[i,2], 
    lon = extraction_points[i,1],
    depth = c(point_series)
  )
  
  bathy <- rbind(bathy, data)
  
  rm(data)
  print(i)
}

# cleanup and formatting
bathy <- bathy %>% filter(!is.na(depth))

m7 <- cbind(m7, bathy)
m7 <- m7[,-c(34:36)]

rm(r_brick) #cleanup
```

```{r CHLOROPHYLL}
## Get existing chlorophyll data
chloro.stmp <- omega_combo %>% dplyr::select(c("kode", "chloro"))
blank <- left_join(high_res, chloro.stmp, by = c("kode"))
blank <- filter(blank, is.na(chloro))

## Get VGPM data for missing days
source('~/Desktop/blues-makos/analyzePSAT/R/vgpm.load.r')
library(raster)

tVec <- c(seq.Date(as.Date('2015-10-01'), as.Date('2016-04-01'), by='month'),
          seq.Date(as.Date('2016-09-01'), as.Date('2017-03-01'), by='month'),
          seq.Date(as.Date('2017-11-01'), as.Date('2018-01-01'), by='month')
)
for (i in 1:length(tVec)){
  doy <- lubridate::yday(tVec[i])
  if (nchar(doy) == 1) doy <- paste('00', doy, sep='')
  if (nchar(doy) == 2) doy <- paste('0', doy, sep='')
  yrday <- paste(lubridate::year(tVec[i]), doy, sep='')
  url <- paste('http://orca.science.oregonstate.edu/data/1x2/monthly/vgpm.r2018.m.chl.m.sst/xyz/vgpm.', yrday, '.all.xyz.gz', sep='')
  filename <- paste('vgpm.', yrday, '.all.xyz.gz', sep='')
  #repeat{
  curl::curl_download(url, filename, quiet=FALSE)
  
  #  tryCatch({
  #    err <- try(RNetCDF::open.nc(filename), silent = T)
  #  }, error=function(e){print(paste('ERROR: Download of data at ', yrday, ' failed. Trying call to server again.', sep = ''))})
  #  if(class(err) != 'try-error') break
  #}
  print(paste(i))
}

fileList <- list.files("~/Desktop/blues-makos/chloro/", full.names = TRUE)
for (i in 1:length(fileList)){
  system(paste('gunzip -k ', fileList[i], sep=''))
}

for (i in 1:17){
  t1 <- Sys.time()
  doy <- lubridate::yday(tVec[i])
  if (nchar(doy) == 1) doy <- paste('00', doy, sep='')
  if (nchar(doy) == 2) doy <- paste('0', doy, sep='')
  yrday <- paste(lubridate::year(tVec[i]), doy, sep='')
  # load vgpm and rasterize
  vgpm <- vgpm.load(file=paste('~/Desktop/blues-makos/chloro/vgpm.', yrday, '.all.xyz', sep=''), w.lon = -180, e.lon = 180,
                    n.lat = 90, s.lat = -90, raster = TRUE)
  vgpm <- vgpm / 1000 # convert from mg to g C/ m^2 / d^1
  #t2 <- Sys.time()
  #print(paste(t2 - t1)) # print how long it took
  crs_geo <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0"
  crs(vgpm) <- crs_geo
  t3 <- Sys.time()
  vgpm_proj <- projectRaster(vgpm, crs=CRS("+init=epsg:3832"))
  
  #plot(vgpm_proj)
  
  e <- extent(250, 330, 0, 50)
  e <- as(e, 'SpatialPolygons')
  crs(e) <- crs_geo
  e <- spTransform(e, CRS("+init=epsg:3832"))
  
  vgpm_new <- mask(vgpm_proj, e)
  vgpm_new <- projectRaster(vgpm_new, crs=crs_geo)
  
  writeRaster(vgpm_new, file=paste('/Volumes/A_DRIVE/Chlorophyll/vgpm.', yrday, '.all.grd', sep=''), overwrite=T)
  rm(vgpm); rm(vgpm_new); rm(vgpm_proj); gc()
  t4 <- Sys.time()
  
  print(paste(i, 'took', round(t4 - t1), 'secs'))
}
#plot(vgpm); world(add=T); plot(e, add=T)

fileList <- list.files("/Volumes/A_DRIVE/Chlorophyll/", full.names = TRUE)
fileList <- fileList[grep('.gri', fileList)]

# extract monthly chlorophyll values for each track location
tVec <- c(seq.Date(as.Date('2015-10-01'), as.Date('2016-04-01'), by='month'),
          seq.Date(as.Date('2016-09-01'), as.Date('2017-03-01'), by='month'),
          seq.Date(as.Date('2017-11-01'), as.Date('2018-02-01'), by='month')
)

data <- data.frame()

for(i in 1:17) {
  doy <- lubridate::yday(tVec[i])
  if (nchar(doy) == 1) doy <- paste('00', doy, sep='')
  if (nchar(doy) == 2) doy <- paste('0', doy, sep='')
  yrday <- paste(lubridate::year(tVec[i]), doy, sep='')
  file <- fileList[grep(yrday, fileList)]
  c_brick <- raster(file)
  sub1 <- blank[blank$Date >= tVec[i] & blank$Date <= tVec[i+1],]
  extraction_points <- data.frame(
    x = sub1$longitude,
    y = sub1$latitude
  )
  for(tt in 1:nrow(extraction_points)) {
    point_series <- raster::extract(c_brick, extraction_points[tt,], method='simple')
    sub1[tt,43] <- point_series
  }
  
  data <- rbind(data, sub1)
  print(i)
}



# transform the array into a raster layer
r_brick <- raster(depth.array, xmn=min(lat), xmx=max(lat), ymn=min(lon), ymx=max(lon), crs=CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs+ towgs84=0,0,0"))
# check orrientation
plot(r_brick)
# correct the orrientation of the raster brick
r_brick <- flip(t(r_brick), direction='y')

# extracting points from the brick
bathy <- data.frame()

for(i in 1:nrow(extraction_points)) {
  point_series <- raster::extract(r_brick, extraction_points[i,], method='simple')
  data <- data.frame(
    point = c(i),
    lat = extraction_points[i,2], 
    lon = extraction_points[i,1],
    depth = c(point_series)
  )
  
  bathy <- rbind(bathy, data)
  
  rm(data)
  print(i)
}

# cleanup and formatting
bathy <- bathy %>% filter(!is.na(depth))

m7 <- cbind(m7, bathy)
m7 <- m7[,-c(34:36)]

rm(r_brick) #cleanup


```

# Downloading data from HYCOM
``` {r HYCOM}
sp.lim <- list(lonmin = -79, lonmax = -10, latmin = 7, latmax = 48)

get.env(uniqueDates = as.POSIXct("2015-11-24"), filename = "hycom", type = "hycom", spatLim = sp.lim, save.dir = '~/Desktop/')
```

# Extracting Temperature at Depth 
```{r T200}
# MAKOS
# STEP 1: Reclaim temperature at depth values from PSAT tags
mako_mod <- omega_combo %>% filter(species == "I. oxyrinchus")

z <- data.frame()

for (i in 1:length(makos)) {
  etuff <- read_etuff(makos[i])
  temp <- get_pdt(etuff)
  temp$ptt <- etuff$meta$ptt
  #temp <- temp %>% separate(DateTime, into = c("Date", "Time"), sep = "([ ])")
  temp$Date <- as.character(temp$Date)
  #temp <- temp[,-c(2)]
  temp <- temp %>% filter(PdtDepth == 200)
  temp <- temp %>% transmute(
    kode = paste(Date, ptt, sep = "_"),
    "Temp200" = ((PdtTempMax + PdtTempMin)/2) 
  )
  alt <- inner_join(mako_mod, temp, by = "kode")
  z <- rbind(z, alt)
  print(i)
  rm(etuff, temp, alt)
}

## Temperature at 200 meters extracted for 34 days
alt <- left_join(mako_mod, z)

# find days that are still missing temperature data
temp <- alt %>% filter(is.na(Temp200))

# STEP 2: Extract missing values from NetCDF files
hList <- list.files(path = "/Volumes/A_DRIVE/HYCOM/",full.names = T)

mdate <- unique(mako_mod$Date.y) 

mtemp <- data.frame()

for (i in 1:length(mdate)) {
  r <- hList[grep(paste(mdate[i]), hList)]
  sub <- temp %>% filter(Date.y == mdate[i])
  r <- raster(r, band = 23)
  extract_pts <- sub[,4:5]
  ext <- raster::extract(r, extract_pts, method = "simple")
  sub$Temp200 <- ext
  mtemp <- rbind(mtemp, sub)
  print(i)
  rm(r,sub,extract_pts,ext)
}
# ~251 temperature values extracted from this code

# separate the blanks
blanks <- mtemp %>% filter(is.na(Temp200))
# separate filled values
mtemp <-  mtemp %>% filter(!is.na(Temp200))

# values are still missing because the longitude coorinate is in the wrong format - this loop will correct it
for (i in 1:nrow(blanks)) {
  r <- hList[grep(paste(blanks[i,1]), hList)]
  sub <- blanks[i,]
  r <- raster(r, band = 23)
  extract_pts <- sub[,4:5]
  extract_pts$x <- extract_pts$x + 360 # correct a different format in longitude
  ext <- raster::extract(r, extract_pts, method = "simple")
  sub$Temp200 <- ext
  mtemp <- rbind(mtemp, sub)
  print(i)
  rm(r,sub,extract_pts,ext)
}

rm(blanks)

# get rid of NA's that have now been filled
alt <- alt %>% filter(!is.na(Temp200))

# combine observed and modelled data
mtemp <- rbind(alt, mtemp)
mako_mod <- mtemp
rm(mtemp)

# BLUE SHARKS
## Extracting data from Depth-Temperature Profiles
blue_mod <- omega_combo %>% filter(species == "P. glauca")

z <- data.frame()

for (i in 1:length(blues)) {
  etuff <- read_etuff(blues[i])
  temp_blue <- get_pdt(etuff)
  temp_blue$ptt <- etuff$meta$ptt
  #temp_blue <- temp_blue %>% separate(DateTime, into = c("Date", "Time"), sep = "([ ])")
  temp_blue$Date <- as.character(temp_blue$Date)
  #temp_blue <- temp_blue[,-c(2)]
  temp_blue <- temp_blue %>% filter(PdtDepth == 200)
  temp_blue <- temp_blue %>% transmute(
    kode = paste(Date, ptt, sep = "_"),
    "Temp200" = ((PdtTempMax + PdtTempMin)/2) 
  )
  alt <- inner_join(blue_mod, temp_blue, by = "kode")
  z <- rbind(z, alt)
  print(i)
  rm(etuff, temp, alt)
}
# Produces 100 observed values for Temp 200

alt_blue <- left_join(blue_mod, z)

temp_blue <- alt_blue %>% filter(is.na(Temp200))

# Using HYCOM data to fill in the blanks

bdate <- unique(blue_mod$Date.y) 

btemp <- data.frame()

for (i in 1:length(bdate)) {
  r <- hList[grep(paste(bdate[i]), hList)]
  sub <- temp_blue %>% filter(Date.y == bdate[i])
  r <- raster(r, band = 23)
  extract_pts <- sub[,4:5]
  extract_pts$x <- extract_pts$x + 360
  ext <- raster::extract(r, extract_pts, method = "simple")
  sub$Temp200 <- ext
  btemp <- rbind(btemp, sub)
  print(i)
  rm(r,sub,extract_pts,ext)
}
# all remaining NAs occur in water more shallow than 200 meters

# get rid of NAs which have now been filled
alt_blue <- alt_blue %>% filter(!is.na(Temp200))
btemp <- rbind(alt_blue, btemp)
blue_mod <- btemp
rm(btemp)
```

## MULTINOMIAL LOGISTIC REGRESSION ANALYSIS ##
# Cross correlation analysis
Determine if any of your environmental variables are colinear
```{r cortest}
combo_mod <- rbind(mako_mod, blue_mod)
combo_mod <- combo_mod %>% filter(cluster != 8 & cluster != 9)
c.cor <- cor(combo_mod[,c(4:15,30)], use = "complete.obs")
```

## FIXED EFFECT MODEL 
# FORWARD SELECTION 1
Identify the best single predictive variable for cluster placement 
``` {r FS1}
combo_mod$clus2 <- relevel(as.factor(combo_mod$cluster), ref = 3)
combo_mod$species <- relevel(as.factor(combo_mod$species), ref = "P. glauca")
combo_mod$ptt <- factor(combo_mod$ptt)
#combo_mod$ptt <- relevel(as.factor(combo_mod$ptt), ref = "106754")

full <- combo_mod %>% filter(lunar >= 0.75)
full$lunar <- 2
mid <- combo_mod %>% filter(lunar < 0.75)
mid$lunar <- 1
combo_mod <- rbind(full, mid)
combo_mod$lunar <- relevel(as.factor(combo_mod$lunar), ref = 1)
rm(full, mid)

combo_mod <- combo_mod %>% separate(Date, into = c("Year", "Month", "Day"), sep = "([-])")
greg <- combo_mod %>% filter(Year == 2008 | Year == 2009)
greg$Year <- 1
cam <- combo_mod %>% filter(Year != 2008 & Year != 2009)
cam$Year <- 2
combo_mod <- rbind(cam, greg)
rm(cam, greg)
combo_mod$Year <- relevel(as.factor(combo_mod$Year), ref = "2")
combo_mod$Month <- relevel(as.factor(combo_mod$Month), ref = "12")


mod01 <- multinom(clus2 ~ sst, data = combo_mod) # AIC: 3740.653
mod02 <- multinom(clus2 ~ sst_sd, data = combo_mod) # AIC: 3921.486
mod03 <- multinom(clus2 ~ ssh, data = combo_mod) # AIC: 3817.16
mod04 <- multinom(clus2 ~ ssh_sd, data = combo_mod) # AIC: 4224.496 
mod05 <- multinom(clus2 ~ eke, data = combo_mod) # AIC: 4242.071 
mod06 <- multinom(clus2 ~ ild.5, data = combo_mod) # AIC: 4235.485
mod07 <- multinom(clus2 ~ n2, data = combo_mod) # AIC: 4207.371
mod08 <- multinom(clus2 ~ bathy, data = combo_mod) # AIC: 4117.645
mod09 <- multinom(clus2 ~ chloro, data = combo_mod) # AIC: 4014.514
mod10 <- multinom(clus2 ~ lunar, data = combo_mod) # AIC: 4236.469 
mod11 <- multinom(clus2 ~ species, data = combo_mod) # AIC: 3819.802
mod12 <- multinom(clus2 ~ x, data = combo_mod) # AIC: 3986.341
mod13 <- multinom(clus2 ~ y, data = combo_mod) # AIC: 3954.994
mod14 <- multinom(clus2 ~ Year, data = combo_mod) # AIC: 3607.569 OR 3807.989 for greg vs. cam
mod15 <- multinom(clus2 ~ Month, data = combo_mod) # AIC: 4139.339

vec_AIC <- AIC(mod01, mod02, mod03, mod04, mod05, mod06, mod07, mod08, mod09, mod10, mod11, mod12, mod13, mod14, mod15)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2))
AICw
# Chlorophyll is best predictor
rm(mod01, mod02, mod03, mod04, mod05, mod06, mod07, mod08, mod10, mod11, mod13, mod14, mod15) # cleanup
# z-test
z <- summary(mod09)$coefficients/summary(mod09)$standard.errors
p <- (1 - pnorm(abs(z), 0, 1)) * 2
```

# FORWARD SELECTION 2
Identify the second best predictive variable for cluster placement
``` {r FS2}
mod1.01 <- multinom(clus2 ~ chloro + sst, data = combo_mod)
mod1.02 <- multinom(clus2 ~ chloro + sst_sd, data = combo_mod)
mod1.03 <- multinom(clus2 ~ chloro + ssh_sd, data = combo_mod)
mod1.04 <- multinom(clus2 ~ chloro + eke, data = combo_mod)
mod1.05 <- multinom(clus2 ~ chloro + ild.5, data = combo_mod)
mod1.06 <- multinom(clus2 ~ chloro + n2, data = combo_mod)
mod1.07 <- multinom(clus2 ~ chloro + lunar, data = combo_mod)
mod1.08 <- multinom(clus2 ~ chloro + species, data = combo_mod)
mod1.09 <- multinom(clus2 ~ chloro + x, data = combo_mod)
mod1.10 <- multinom(clus2 ~ chloro + y, data = combo_mod)
mod1.11 <- multinom(clus2 ~ chloro + Year, data = combo_mod)
mod1.12 <- multinom(clus2 ~ chloro + Month, data = combo_mod)

vec_AIC <- AIC(mod1.01, mod1.02, mod1.03, mod1.04, mod1.05, mod1.06, mod1.07, mod1.08, mod1.09, mod1.10, mod1.11, mod1.12)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2))
AICw
# Year is best predictor
rm(mod1.01, mod1.02, mod1.03, mod1.04, mod1.05, mod1.06, mod1.07, mod1.08, mod1.10, mod1.12) # cleanup

# z-test
z <- summary(mod1.11)$coefficients/summary(mod1.11)$standard.errors
p <- (1 - pnorm(abs(z), 0, 1)) * 2
```

# FORWARD SELECTION 3
Identify the third best predictive variable for cluster placement
``` {r FS3}
mod1.1.01 <- multinom(clus2 ~ chloro + Year + sst, data = combo_mod)
mod1.1.02 <- multinom(clus2 ~ chloro + Year + sst_sd, data = combo_mod)
mod1.1.03 <- multinom(clus2 ~ chloro + Year + ssh_sd, data = combo_mod)
mod1.1.04 <- multinom(clus2 ~ chloro + Year + eke, data = combo_mod)
mod1.1.05 <- multinom(clus2 ~ chloro + Year + ild.5, data = combo_mod)
mod1.1.06 <- multinom(clus2 ~ chloro + Year + n2, data = combo_mod)
mod1.1.07 <- multinom(clus2 ~ chloro + Year + lunar, data = combo_mod)
mod1.1.08 <- multinom(clus2 ~ chloro + Year + species, data = combo_mod)
mod1.1.09 <- multinom(clus2 ~ chloro + Year + x, data = combo_mod)
mod1.1.10 <- multinom(clus2 ~ chloro + Year + y, data = combo_mod)
mod1.1.11 <- multinom(clus2 ~ chloro + Year + Month, data = combo_mod)

vec_AIC <- AIC(mod1.1.01, mod1.1.02, mod1.1.03, mod1.1.04, mod1.1.05, mod1.1.06, mod1.1.07, mod1.1.08, mod1.1.09, mod1.1.10, mod1.1.11)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2))
AICw
# Species is the next best predictor
rm(mod1.1.01, mod1.1.02, mod1.1.03, mod1.1.04, mod1.1.05, mod1.1.06, mod1.1.07, mod1.1.09, mod1.1.10, mod1.1.11) # cleanup

# z-test
z <- summary(mod1.1.08)$coefficients/summary(mod1.1.08)$standard.errors
p <- (1 - pnorm(abs(z), 0, 1)) * 2
```

# FORWARD SELECTION 4
Identify the fourth best predictive variable for cluster placement
```{r FS4}
mod1.1.1.01 <- multinom(clus2 ~ chloro + Year + species + sst, data = combo_mod)
mod1.1.1.02 <- multinom(clus2 ~ chloro + Year + species + sst_sd, data = combo_mod)
mod1.1.1.03 <- multinom(clus2 ~ chloro + Year + species + ssh_sd, data = combo_mod)
mod1.1.1.04 <- multinom(clus2 ~ chloro + Year + species + eke, data = combo_mod)
mod1.1.1.05 <- multinom(clus2 ~ chloro + Year + species + ild.5, data = combo_mod)
mod1.1.1.06 <- multinom(clus2 ~ chloro + Year + species + n2, data = combo_mod)
mod1.1.1.07 <- multinom(clus2 ~ chloro + Year + species + lunar, data = combo_mod)
mod1.1.1.08 <- multinom(clus2 ~ chloro + Year + species + x, data = combo_mod)
mod1.1.1.09 <- multinom(clus2 ~ chloro + Year + species + y, data = combo_mod)
mod1.1.1.10 <- multinom(clus2 ~ chloro + Year + species + Month, data = combo_mod)

vec_AIC <- AIC(mod1.1.08, mod1.1.1.01, mod1.1.1.02, mod1.1.1.03, mod1.1.1.04, mod1.1.1.05, mod1.1.1.06, mod1.1.1.07, mod1.1.1.08, mod1.1.1.09, mod1.1.1.10)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2))
AICw
# Latitude is best
rm(mod1.1.1.01, mod1.1.1.02, mod1.1.1.03, mod1.1.1.04, mod1.1.1.05, mod1.1.1.06, mod1.1.1.07, mod1.1.1.08, mod1.1.1.10)

# z-test
z <- summary(mod1.1.1.09)$coefficients/summary(mod1.1.1.09)$standard.errors
p <- (1 - pnorm(abs(z), 0, 1)) * 2
```

# FORWARD SELECTION 5
Identify the fifth best predictive variable for cluster placement
```{r FS5}
mod1.1.1.1.01 <- multinom(clus2 ~ chloro + Year + species + y + sst_sd, data = combo_mod) 
mod1.1.1.1.02 <- multinom(clus2 ~ chloro + Year + species + y + ssh_sd, data = combo_mod)
mod1.1.1.1.03 <- multinom(clus2 ~ chloro + Year + species + y + eke, data = combo_mod)
mod1.1.1.1.04 <- multinom(clus2 ~ chloro + Year + species + y + ild.5, data = combo_mod)
mod1.1.1.1.05 <- multinom(clus2 ~ chloro + Year + species + y + n2, data = combo_mod)
mod1.1.1.1.06 <- multinom(clus2 ~ chloro + Year + species + y + lunar, data = combo_mod)
mod1.1.1.1.07 <- multinom(clus2 ~ chloro + Year + species + y + x, data = combo_mod)
mod1.1.1.1.08 <- multinom(clus2 ~ chloro + Year + species + y + Month, data = combo_mod)
  
vec_AIC <- AIC(mod1.1.1.09, mod1.1.1.1.01, mod1.1.1.1.02, mod1.1.1.1.03, mod1.1.1.1.04, mod1.1.1.1.05, mod1.1.1.1.06, mod1.1.1.1.07, mod1.1.1.1.08)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2))
AICw
# longitude is best
rm(mod1.1.1.1.01, mod1.1.1.1.02, mod1.1.1.1.03, mod1.1.1.1.04, mod1.1.1.1.05, mod1.1.1.1.06, mod1.1.1.1.08)

# z-test
z <- summary(mod1.1.1.1.07)$coefficients/summary(mod1.1.1.1.07)$standard.errors
p <- (1 - pnorm(abs(z), 0, 1)) * 2
```

# FORWARD SELECTION 6
Identify the sixth best predictive variable for cluster placement
``` {r FS6}
mod1.1.1.1.1.01 <- multinom(clus2 ~ chloro + Year + species + y + x + sst_sd, data = combo_mod)
mod1.1.1.1.1.02 <- multinom(clus2 ~ chloro + Year + species + y + x + ssh_sd, data = combo_mod)
mod1.1.1.1.1.03 <- multinom(clus2 ~ chloro + Year + species + y + x + eke, data = combo_mod)
mod1.1.1.1.1.04 <- multinom(clus2 ~ chloro + Year + species + y + x + ild.5, data = combo_mod)
mod1.1.1.1.1.05 <- multinom(clus2 ~ chloro + Year + species + y + x + n2, data = combo_mod)
mod1.1.1.1.1.06 <- multinom(clus2 ~ chloro + Year + species + y + x + lunar, data = combo_mod)
mod1.1.1.1.1.07 <- multinom(clus2 ~ chloro + Year + species + y + x + Month, data = combo_mod)

vec_AIC <- AIC(mod1.1.1.1.07, mod1.1.1.1.1.01, mod1.1.1.1.1.02, mod1.1.1.1.1.03, mod1.1.1.1.1.04, mod1.1.1.1.1.05, mod1.1.1.1.1.06, mod1.1.1.1.1.07)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2))
AICw
# Lunar illumination is the next best predictor
rm(mod1.1.1.1.1.01, mod1.1.1.1.1.02, mod1.1.1.1.1.03, mod1.1.1.1.1.04, mod1.1.1.1.1.05, mod1.1.1.1.1.07)

z <- summary(mod1.1.1.1.1.06)$coefficients/summary(mod1.1.1.1.1.06)$standard.errors
p <- (1 - pnorm(abs(z), 0, 1)) * 2
```

# FORWARD SELECTION 7
Identify the seventh best predictive variable for cluster placement
``` {r FS7}
mod1.1.1.1.1.1.01 <- multinom(clus2 ~ chloro + Year + species + y + x + lunar + sst_sd, data = combo_mod)
mod1.1.1.1.1.1.02 <- multinom(clus2 ~ chloro + Year + species + y + x + lunar + ssh_sd, data = combo_mod)
mod1.1.1.1.1.1.03 <- multinom(clus2 ~ chloro + Year + species + y + x + lunar + eke, data = combo_mod)
mod1.1.1.1.1.1.04 <- multinom(clus2 ~ chloro + Year + species + y + x + lunar + ild.5, data = combo_mod)
mod1.1.1.1.1.1.05 <- multinom(clus2 ~ chloro + Year + species + y + x + lunar + n2, data = combo_mod)

vec_AIC <- AIC(mod1.1.1.1.1.06, mod1.1.1.1.1.1.01, mod1.1.1.1.1.1.02, mod1.1.1.1.1.1.03, mod1.1.1.1.1.1.04, mod1.1.1.1.1.1.05)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2))
AICw
# n2 is the next best predictor
rm(mod1.1.1.1.1.1.01, mod1.1.1.1.1.1.02, mod1.1.1.1.1.1.03, mod1.1.1.1.1.1.04)

# z-test
z <- summary(mod1.1.1.1.1.1.05)$coefficients/summary(mod1.1.1.1.1.1.05)$standard.errors
p <- (1 - pnorm(abs(z), 0, 1)) * 2
```

# FORWARD SELECTION 8
Identify the eighth best predictive variable for cluster placement
``` {r FS8}
mod1.1.1.1.1.1.1.01 <- multinom(clus2 ~ chloro + Year + species + y + x + lunar + n2 + sst_sd, data = combo_mod)
mod1.1.1.1.1.1.1.02 <- multinom(clus2 ~ chloro + Year + species + y + x + lunar + n2 + ssh_sd, data = combo_mod)
mod1.1.1.1.1.1.1.03 <- multinom(clus2 ~ chloro + Year + species + y + x + lunar + n2 + eke, data = combo_mod)

vec_AIC <- AIC(mod1.1.1.1.1.1.05, mod1.1.1.1.1.1.1.01, mod1.1.1.1.1.1.1.02, mod1.1.1.1.1.1.1.03)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2))
AICw
# sea-surface temperature standard deviation is the next best predictor
rm(mod1.1.1.1.1.1.1.02, mod1.1.1.1.1.1.1.03)

# z-test
z <- summary(mod1.1.1.1.1.1.1.01)$coefficients/summary(mod1.1.1.1.1.1.1.01)$standard.errors
p <- (1 - pnorm(abs(z), 0, 1)) * 2
```

# FORWARD SELECTION 9
Identify the ninth best predictive variable for cluster placement
``` {r FS9}
mod1.1.1.1.1.1.1.1.01 <- multinom(clus2 ~ chloro + Year + species + y + x + lunar + n2 + sst_sd + ssh_sd, data = combo_mod)
mod1.1.1.1.1.1.1.1.02 <- multinom(clus2 ~ chloro + Year + species + y + x + lunar + n2 + sst_sd + eke, data = combo_mod)

vec_AIC <- AIC(mod1.1.1.1.1.1.1.01, mod1.1.1.1.1.1.1.1.01, mod1.1.1.1.1.1.1.1.02)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2))
AICw
# eke is the next best predictor
rm(mod1.1.1.1.1.1.1.1.01)

# z-test
z <- summary(mod1.1.1.1.1.1.1.1.02)$coefficients/summary(mod1.1.1.1.1.1.1.1.02)$standard.errors
p <- (1 - pnorm(abs(z), 0, 1)) * 2
```

# FORWARD SELECTION 10
Identify the tenth best predictive variable for cluster placement
``` {r FS10}
mod1.1.1.1.1.1.1.1.1.01 <- multinom(clus2 ~ chloro + Year + species + y + x + lunar + n2 + sst_sd + eke + ssh_sd, data = combo_mod)

vec_AIC <- AIC(mod1.1.1.1.1.1.1.1.02, mod1.1.1.1.1.1.1.1.1.01)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2))
AICw

# the previous model was better
```

## Including Interaction Terms
Test the interactions between model variables, and their influence on the predictive capacity of the model 
```{r Interactions}
# 1: chloro:Year
mod0.1 <- multinom(clus2 ~ chloro + Year + species + y + x + lunar + n2 + sst_sd + eke + chloro:Year, data = combo_mod)
vec_AIC <- AIC(mod1.1.1.1.1.1.1.1.02, mod0.1)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2)) 
AICw
# this interaction made the model worse

# 2: chloro:species
mod0.1 <- multinom(clus2 ~ chloro + Year + species + y + x + lunar + n2 + sst_sd + eke + chloro:species, data = combo_mod)
vec_AIC <- AIC(mod1.1.1.1.1.1.1.1.02, mod0.1)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2)) 
AICw
# This made the model better
z <- summary(mod0.1)$coefficients/summary(mod0.1)$standard.errors
p <- (1 - pnorm(abs(z), 0, 1)) * 2

# 3: chloro:y
mod0.2 <- multinom(clus2 ~ chloro + Year + species + y + x + lunar + n2 + sst_sd + eke + chloro:species + chloro:y, data = combo_mod)
vec_AIC <- AIC(mod0.1, mod0.2)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2)) 
AICw
# This made the model better
z <- summary(mod0.2)$coefficients/summary(mod0.2)$standard.errors
p <- (1 - pnorm(abs(z), 0, 1)) * 2

# 4: chloro:x
mod0.3 <- multinom(clus2 ~ chloro + Year + species + y + x + lunar + n2 + sst_sd + eke + chloro:species + chloro:y + chloro:x, data = combo_mod)
vec_AIC <- AIC(mod0.2, mod0.3)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2)) 
AICw
# this term made the model worse

# 5: chloro:lunar
mod0.3 <- multinom(clus2 ~ chloro + Year + species + y + x + lunar + n2 + sst_sd + eke + chloro:species + chloro:y + chloro:lunar, data = combo_mod)
vec_AIC <- AIC(mod0.2, mod0.3)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2)) 
AICw
# 50/50 split - go for parsimony

# 6: chloro:n2
mod0.3 <- multinom(clus2 ~ chloro + Year + species + y + x + lunar + n2 + sst_sd + eke + chloro:species + chloro:y + chloro:n2, data = combo_mod)
vec_AIC <- AIC(mod0.2, mod0.3)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2)) 
AICw
# this term made the model worse

# 7: chloro:sst_sd
mod0.3 <- multinom(clus2 ~ chloro + Year + species + y + x + lunar + n2 + sst_sd + eke + chloro:species + chloro:y + chloro:sst_sd, data = combo_mod)
vec_AIC <- AIC(mod0.2, mod0.3)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2)) 
AICw
# this term made the model worse

# 8: chloro:eke
mod0.3 <- multinom(clus2 ~ chloro + Year + species + y + x + lunar + n2 + sst_sd + eke + chloro:species + chloro:y + chloro:eke, data = combo_mod)
vec_AIC <- AIC(mod0.2, mod0.3)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2)) 
AICw
# this term made the model worse

# 9: Year:species
mod0.3 <- multinom(clus2 ~ chloro + Year + species + y + x + lunar + n2 + sst_sd + eke + chloro:species + chloro:y + Year:species, data = combo_mod)
vec_AIC <- AIC(mod0.2, mod0.3)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2)) 
AICw
# 50/50 split

# 10: Year:y
mod0.3 <- multinom(clus2 ~ chloro + Year + species + y + x + lunar + n2 + sst_sd + eke + chloro:species + chloro:y + Year:y, data = combo_mod)
vec_AIC <- AIC(mod0.2, mod0.3)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2)) 
AICw
# 50/50 split

# 11: Year:x
mod0.3 <- multinom(clus2 ~ chloro + Year + species + y + x + lunar + n2 + sst_sd + eke + chloro:species + chloro:y + Year:x, data = combo_mod)
vec_AIC <- AIC(mod0.2, mod0.3)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2)) 
AICw
# this term made the model better
z <- summary(mod0.3)$coefficients/summary(mod0.3)$standard.errors
p <- (1 - pnorm(abs(z), 0, 1)) * 2

# 12: Year:lunar
mod0.4 <- multinom(clus2 ~ chloro + Year + species + y + x + lunar + n2 + sst_sd + eke + chloro:species + chloro:y + Year:x + Year:lunar, data = combo_mod)
vec_AIC <- AIC(mod0.3, mod0.4)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2)) 
AICw
# this term made the model worse

# 13: Year:n2
mod0.4 <- multinom(clus2 ~ chloro + Year + species + y + x + lunar + n2 + sst_sd + eke + chloro:species + chloro:y + Year:x + Year:n2, data = combo_mod)
vec_AIC <- AIC(mod0.3, mod0.4)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2)) 
AICw
# this term made the model worse

# 13.5: Year:sst_sd
mod0.4 <- multinom(clus2 ~ chloro + Year + species + y + x + lunar + n2 + sst_sd + eke + chloro:species + chloro:y + Year:x + Year:sst_sd, data = combo_mod)
vec_AIC <- AIC(mod0.3, mod0.4)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2)) 
AICw
# this term made the model worse

# 14: Year:eke
mod0.4 <- multinom(clus2 ~ chloro + Year + species + y + x + lunar + n2 + sst_sd + eke + chloro:species + chloro:y + Year:x + Year:eke, data = combo_mod)
vec_AIC <- AIC(mod0.3, mod0.4)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2)) 
AICw
# this term made the model worse

# 15: y:x
mod0.4 <- multinom(clus2 ~ chloro + Year + species + y + x + lunar + n2 + sst_sd + eke + chloro:species + chloro:y + Year:x + Year:lunar + y:x, data = combo_mod)
vec_AIC <- AIC(mod0.3, mod0.4)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2)) 
AICw
# this term made the model better
z <- summary(mod0.4)$coefficients/summary(mod0.4)$standard.errors
p <- (1 - pnorm(abs(z), 0, 1)) * 2
# all terms still statistically significant

#16: y:lunar
mod0.5 <- multinom(clus2 ~ chloro + Year + species + y + x + lunar + n2 + sst_sd + eke + chloro:species + chloro:y + Year:x + Year:lunar + y:x + y:lunar, data = combo_mod)
vec_AIC <- AIC(mod0.4, mod0.5)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2)) 
AICw
# this term made the model worse

# 17: y:n2
mod0.5 <- multinom(clus2 ~ chloro + Year + species + y + x + lunar + n2 + sst_sd + eke + chloro:species + chloro:y + Year:x + Year:lunar + y:x + y:n2, data = combo_mod)
vec_AIC <- AIC(mod0.4, mod0.5)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2)) 
AICw
# 50/50 split

#18: y:sst_sd
mod0.5 <- multinom(clus2 ~ chloro + Year + species + y + x + lunar + n2 + sst_sd + eke + chloro:species + chloro:y + Year:x + Year:lunar + y:x + y:sst_sd, data = combo_mod)
vec_AIC <- AIC(mod0.4, mod0.5)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2)) 
AICw
# this term made the model worse

#19: y:eke
mod0.5 <- multinom(clus2 ~ chloro + Year + species + y + x + lunar + n2 + sst_sd + eke + chloro:species + chloro:y + Year:x + Year:lunar + y:x + y:eke, data = combo_mod)
vec_AIC <- AIC(mod0.4, mod0.5)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2)) 
AICw
# this made the model worse

# 20: x:lunar
mod0.5 <- multinom(clus2 ~ chloro + Year + species + y + x + lunar + n2 + sst_sd + eke + chloro:species + chloro:y + Year:x + Year:lunar + y:x + x:lunar, data = combo_mod)
vec_AIC <- AIC(mod0.4, mod0.5)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2)) 
AICw
# this term made the model worse

#21: x:n2
mod0.5 <- multinom(clus2 ~ chloro + Year + species + y + x + lunar + n2 + sst_sd + eke + chloro:species + chloro:y + Year:x + Year:lunar + y:x + x:n2, data = combo_mod)
vec_AIC <- AIC(mod0.4, mod0.5)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2)) 
AICw
# this term made the model better
z <- summary(mod0.5)$coefficients/summary(mod0.5)$standard.errors
p <- (1 - pnorm(abs(z), 0, 1)) * 2
# all terms still significant

#22: x:sst_sd
mod0.6 <- multinom(clus2 ~ chloro + Year + species + y + x + lunar + n2 + sst_sd + eke + chloro:species + chloro:y + Year:x + Year:lunar + y:x + x:n2 + x:sst_sd, data = combo_mod)
vec_AIC <- AIC(mod0.5, mod0.6)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2)) 
AICw
# this term made the model worse

#23: x:eke
mod0.6 <- multinom(clus2 ~ chloro + Year + species + y + x + lunar + n2 + sst_sd + eke + chloro:species + chloro:y + Year:x + Year:lunar + y:x + x:n2 + x:eke, data = combo_mod)
vec_AIC <- AIC(mod0.5, mod0.6)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2)) 
AICw
# this term made the model worse

# 24: lunar:n2
mod0.6 <- multinom(clus2 ~ chloro + Year + species + y + x + lunar + n2 + sst_sd + eke + chloro:species + chloro:y + Year:x + Year:lunar + y:x + x:n2 + lunar:n2, data = combo_mod)
vec_AIC <- AIC(mod0.5, mod0.6)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2)) 
AICw
# this term made the model worse

# 25: lunar:sst_sd
mod0.6 <- multinom(clus2 ~ chloro + Year + species + y + x + lunar + n2 + sst_sd + eke + chloro:species + chloro:y + Year:x + Year:lunar + y:x + x:n2 + lunar:sst_sd, data = combo_mod)
vec_AIC <- AIC(mod0.5, mod0.6)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2)) 
AICw
# this term made the model worse

# 26: lunar:eke
mod0.6 <- multinom(clus2 ~ chloro + Year + species + y + x + lunar + n2 + sst_sd + eke + chloro:species + chloro:y + Year:x + Year:lunar + y:x + x:n2 + lunar:eke, data = combo_mod)
vec_AIC <- AIC(mod0.5, mod0.6)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2)) 
AICw
# this term made the model better
z <- summary(mod0.6)$coefficients/summary(mod0.6)$standard.errors
p <- (1 - pnorm(abs(z), 0, 1)) * 2
# all terms still significant

# 27: n2:sst_sd
mod0.7 <- multinom(clus2 ~ chloro + Year + species + y + x + lunar + n2 + sst_sd + eke + chloro:species + chloro:y + Year:x + Year:lunar + y:x + x:n2 + lunar:eke + n2:sst_sd, data = combo_mod)
vec_AIC <- AIC(mod0.6, mod0.7)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2)) 
AICw
# this made the model worse

# 28: n2:eke
mod0.7 <- multinom(clus2 ~ chloro + Year + species + y + x + lunar + n2 + sst_sd + eke + chloro:species + chloro:y + Year:x + Year:lunar + y:x + x:n2 + lunar:eke + n2:eke, data = combo_mod)
vec_AIC <- AIC(mod0.6, mod0.7)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2)) 
AICw
# this made the model worse

#29: sst_sd:eke
mod0.7 <- multinom(clus2 ~ chloro + Year + species + y + x + lunar + n2 + sst_sd + eke + chloro:species + chloro:y + Year:x + Year:lunar + y:x + x:n2 + lunar:eke + sst_sd:eke, data = combo_mod)
vec_AIC <- AIC(mod0.6, mod0.7)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2)) 
AICw
# this made the model worse

#30: species:y
mod0.7 <- multinom(clus2 ~ chloro + Year + species + y + x + lunar + n2 + sst_sd + eke + chloro:species + chloro:y + Year:x + Year:lunar + y:x + x:n2 + lunar:eke + sst_sd:eke + species:y, data = combo_mod)
vec_AIC <- AIC(mod0.6, mod0.7)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2)) 
AICw
# this made the model worse

#31 species:x
mod0.7 <- multinom(clus2 ~ chloro + Year + species + y + x + lunar + n2 + sst_sd + eke + chloro:species + chloro:y + Year:x + Year:lunar + y:x + x:n2 + lunar:eke + sst_sd:eke + species:x, data = combo_mod)
vec_AIC <- AIC(mod0.6, mod0.7)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2)) 
AICw
# this made the model worse

# 32: species:lunar
mod0.7 <- multinom(clus2 ~ chloro + Year + species + y + x + lunar + n2 + sst_sd + eke + chloro:species + chloro:y + Year:x + Year:lunar + y:x + x:n2 + lunar:eke + sst_sd:eke + species:lunar, data = combo_mod)
vec_AIC <- AIC(mod0.6, mod0.7)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2)) 
AICw
# this made the model better
z <- summary(mod0.7)$coefficients/summary(mod0.7)$standard.errors
p <- (1 - pnorm(abs(z), 0, 1)) * 2

# 33: species:n2
mod0.8 <- multinom(clus2 ~ chloro + Year + species + y + x + lunar + n2 + sst_sd + eke + chloro:species + chloro:y + Year:x + Year:lunar + y:x + x:n2 + lunar:eke + sst_sd:eke + species:lunar + species:n2, data = combo_mod)
vec_AIC <- AIC(mod0.7, mod0.8)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2)) 
AICw
# this term made the model worse

#34: species:sst_sd
mod0.8 <- multinom(clus2 ~ chloro + Year + species + y + x + lunar + n2 + sst_sd + eke + chloro:species + chloro:y + Year:x + Year:lunar + y:x + x:n2 + lunar:eke + sst_sd:eke + species:lunar + species:sst_sd, data = combo_mod)
vec_AIC <- AIC(mod0.7, mod0.8)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2)) 
AICw
# 50/50

#35: species:eke
mod0.8 <- multinom(clus2 ~ chloro + Year + species + y + x + lunar + n2 + sst_sd + eke + chloro:species + chloro:y + Year:x + Year:lunar + y:x + x:n2 + lunar:eke + sst_sd:eke + species:lunar + species:eke, data = combo_mod)
vec_AIC <- AIC(mod0.7, mod0.8)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2)) 
AICw
# 50/50

```

# The final, fixed effect multinomial logistic regression model predicting blue and mako shark cluster placement is: mod0.7 

- Chlorophyll
- Year
- Species
- Latitude
- Longitude
- Lunar illumination
- Buoyancy frequency
- Sea-surface temperature standard deviation
- Eddy kinetic energy

- Chlorophyll : Species
- Chlorophyll : Latitude
- Year : Longitude
- Year : Lunar illumination
- Species : Lunar illumination
- Latitude : Longitude
- Longitude : Buoyancy-frequency
- Lunar illumination : Eddy kinetic energy
- Sea-surface standard deviation : Eddy kinetic energy

## MIXED EFFECTS MODEL
# FORWARD SELECTION 1
Identify the best single predictive variable for cluster placement 
``` {r MFS1}
m.mod01 <- mblogit(clus2 ~ sst, data = combo_mod)
m.mod02 <- mblogit(clus2 ~ sst_sd, data = combo_mod)
m.mod03 <- mblogit(clus2 ~ ssh, data = combo_mod)
m.mod04 <- mblogit(clus2 ~ ssh_sd, data = combo_mod)
m.mod05 <- mblogit(clus2 ~ eke, data = combo_mod)
m.mod06 <- mblogit(clus2 ~ ild.5, data = combo_mod)
m.mod07 <- mblogit(clus2 ~ n2, data = combo_mod)
m.mod08 <- mblogit(clus2 ~ chloro, data = combo_mod)
m.mod09 <- mblogit(clus2 ~ lunar, data = combo_mod)
m.mod10 <- mblogit(clus2 ~ bathy, data = combo_mod)
m.mod11 <- mblogit(clus2 ~ 1, random = ~1|species, data = combo_mod)
mblogit(cluster ~ 1, random = ~1|species, data = df)
m.mod12 <- mblogit(clus2 ~ x, data = combo_mod)
m.mod13 <- mblogit(clus2 ~ y, data = combo_mod)

vec_AIC <- AIC(m.mod01, m.mod02, m.mod03, m.mod04, m.mod05, m.mod06, m.mod07, m.mod08, m.mod09, m.mod10, m.mod11, m.mod12, m.mod13)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2))
AICw
# Species is best predictor
rm(m.mod01, m.mod02, m.mod03, m.mod04, m.mod05, m.mod06, m.mod07, m.mod08, m.mod09, m.mod10, m.mod12, m.mod13) # cleanup
```

# FORWARD SELECTION 2
Identify the second best single predictive variable for cluster placement 
``` {r MFS2}
m.mod01.1 <- mblogit(clus2 ~ species + sst, random = ~1|ptt, data = combo_mod)
m.mod01.2 <- mblogit(clus2 ~ species + sst_sd, random = ~1|ptt, data = combo_mod)
m.mod01.3 <- mblogit(clus2 ~ species + ssh, random = ~1|ptt, data = combo_mod)
m.mod01.4 <- mblogit(clus2 ~ species + ssh_sd, random = ~1|ptt, data = combo_mod)
m.mod01.5 <- mblogit(clus2 ~ species + eke, random = ~1|ptt, data = combo_mod)
m.mod01.6 <- mblogit(clus2 ~ species + ild.5, random = ~1|ptt, data = combo_mod)
m.mod01.7 <- mblogit(clus2 ~ species + n2, random = ~1|ptt, data = combo_mod)
m.mod01.8 <- mblogit(clus2 ~ species + bathy, random = ~1|ptt, data = combo_mod)
m.mod01.9 <- mblogit(clus2 ~ species + chloro, random = ~1|ptt, data = combo_mod)
m.mod01.10 <- mblogit(clus2 ~ species + lunar, random = ~1|ptt, data = combo_mod)
m.mod01.11 <- mblogit(clus2 ~ species + x, random = ~1|ptt, data = combo_mod)
m.mod01.12 <- mblogit(clus2 ~ species + y, random = ~1|ptt, data = combo_mod)

vec_AIC <- AIC(m.mod01.1, m.mod01.2, m.mod01.3, m.mod01.4, m.mod01.5, m.mod01.6, m.mod01.7, m.mod01.8, m.mod01.9, m.mod01.10, m.mod01.11, m.mod01.12)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2))
AICw
# Chlorophyll is best predictor
rm(m.mod01.1, m.mod01.2, m.mod01.3, m.mod01.4, m.mod01.5, m.mod01.6, m.mod01.7, m.mod01.8, m.mod01.10, m.mod01.11, m.mod01.12) # cleanup
```

# FORWARD SELECTION 3
Identify the third best single predictive variable for cluster placement 
``` {r MFS3}
m.mod1.1.01 <- mblogit(clus2 ~ species + chloro + sst_sd, random = ~1|ptt, data = combo_mod)
m.mod1.1.02 <- mblogit(clus2 ~ species + chloro + ssh, random = ~1|ptt, data = combo_mod)
m.mod1.1.03 <- mblogit(clus2 ~ species + chloro + ssh_sd, random = ~1|ptt, data = combo_mod)
m.mod1.1.04 <- mblogit(clus2 ~ species + chloro + eke, random = ~1|ptt, data = combo_mod)
m.mod1.1.05 <- mblogit(clus2 ~ species + chloro + ild.5, random = ~1|ptt, data = combo_mod)
m.mod1.1.06 <- mblogit(clus2 ~ species + chloro + n2, random = ~1|ptt, data = combo_mod)
m.mod1.1.07 <- mblogit(clus2 ~ species + chloro + lunar, random = ~1|ptt, data = combo_mod)
m.mod1.1.08 <- mblogit(clus2 ~ species + chloro + x, random = ~1|ptt, data = combo_mod)
m.mod1.1.09 <- mblogit(clus2 ~ species + chloro + y, random = ~1|ptt, data = combo_mod)

vec_AIC <- AIC(m.mod1.1.01, m.mod1.1.02, m.mod1.1.03, m.mod1.1.04, m.mod1.1.05, m.mod1.1.06, m.mod1.1.07, m.mod1.1.08, m.mod1.1.09)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2))
AICw
# Latitude is best predictor
rm(m.mod1.1.01, m.mod1.1.02, m.mod1.1.03, m.mod1.1.04, m.mod1.1.05, m.mod1.1.06, m.mod1.1.07, m.mod1.1.08) # cleanup
```

# FORWARD SELECTION 4
Identify the fourth best single predictive variable for cluster placement 
``` {r MFS4}
m.mod1.1.1.01 <- mblogit(clus2 ~ species + chloro + y + sst_sd, random = ~1|ptt, data = combo_mod)
m.mod1.1.1.03 <- mblogit(clus2 ~ species + chloro + y + ssh_sd, random = ~1|ptt, data = combo_mod)
m.mod1.1.1.04 <- mblogit(clus2 ~ species + chloro + y + eke, random = ~1|ptt, data = combo_mod)
m.mod1.1.1.05 <- mblogit(clus2 ~ species + chloro + y + ild.5, random = ~1|ptt, data = combo_mod)
m.mod1.1.1.06 <- mblogit(clus2 ~ species + chloro + y + n2, random = ~1|ptt, data = combo_mod)
m.mod1.1.1.07 <- mblogit(clus2 ~ species + chloro + y + lunar, random = ~1|ptt, data = combo_mod)
m.mod1.1.1.08 <- mblogit(clus2 ~ species + chloro + y + x, random = ~1|ptt, data = combo_mod)

vec_AIC <- AIC(m.mod1.1.1.01, m.mod1.1.1.03, m.mod1.1.1.04, m.mod1.1.1.05, m.mod1.1.1.06, m.mod1.1.1.08)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2))
AICw
# Lunar illumination is best predictor
rm(m.mod1.1.1.01, m.mod1.1.1.03, m.mod1.1.1.04, m.mod1.1.1.05, m.mod1.1.1.06, m.mod1.1.1.08) # cleanup
```

# FORWARD SELECTION 5
Identify the fifth best single predictive variable for cluster placement 
``` {r MFS5}
m.mod1.1.1.1.01 <- mblogit(clus2 ~ species + chloro + y + lunar + sst_sd, random = ~1|ptt, data = combo_mod)
m.mod1.1.1.1.02 <- mblogit(clus2 ~ species + chloro + y + lunar + ssh_sd, random = ~1|ptt, data = combo_mod)
m.mod1.1.1.1.03 <- mblogit(clus2 ~ species + chloro + y + lunar + eke, random = ~1|ptt, data = combo_mod)
m.mod1.1.1.1.04 <- mblogit(clus2 ~ species + chloro + y + lunar + ild.5, random = ~1|ptt, data = combo_mod)
m.mod1.1.1.1.05 <- mblogit(clus2 ~ species + chloro + y + lunar + n2, random = ~1|ptt, data = combo_mod)
m.mod1.1.1.1.06 <- mblogit(clus2 ~ species + chloro + y + lunar + x, random = ~1|ptt, data = combo_mod)


vec_AIC <- AIC(m.mod1.1.1.07, m.mod1.1.1.1.01, m.mod1.1.1.1.02, m.mod1.1.1.1.03, m.mod1.1.1.1.04, m.mod1.1.1.1.05, m.mod1.1.1.1.06)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2))
AICw
# n2 is best predictor
rm(m.mod1.1.1.1.01, m.mod1.1.1.1.02, m.mod1.1.1.1.03, m.mod1.1.1.1.04, m.mod1.1.1.1.06) # cleanup
```

# FORWARD SELECTION 6
Identify the sixth best single predictive variable for cluster placement 
``` {r MFS6}
m.mod1.1.1.1.1.01 <- mblogit(clus2 ~ species + chloro + y + lunar + n2 + sst_sd, random = ~1|ptt, data = combo_mod)
m.mod1.1.1.1.1.02 <- mblogit(clus2 ~ species + chloro + y + lunar + n2 + ssh_sd, random = ~1|ptt, data = combo_mod)
m.mod1.1.1.1.1.03 <- mblogit(clus2 ~ species + chloro + y + lunar + n2 + eke, random = ~1|ptt, data = combo_mod)
m.mod1.1.1.1.1.04 <- mblogit(clus2 ~ species + chloro + y + lunar + n2 + x, random = ~1|ptt, data = combo_mod)


vec_AIC <- AIC(m.mod1.1.1.1.05, m.mod1.1.1.1.1.01, m.mod1.1.1.1.1.02, m.mod1.1.1.1.1.03, m.mod1.1.1.1.1.04)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2))
AICw
# eke is best predictor
rm(m.mod1.1.1.1.1.01, m.mod1.1.1.1.1.02, m.mod1.1.1.1.1.04) # cleanup
```

# FORWARD SELECTION 7
Identify the seventh best single predictive variable for cluster placement 
``` {r MFS7}
m.mod1.1.1.1.1.1.01 <- mblogit(clus2 ~ species + chloro + y + lunar + n2 + eke + sst_sd, random = ~1|ptt, data = combo_mod)
m.mod1.1.1.1.1.1.02 <- mblogit(clus2 ~ species + chloro + y + lunar + n2 + eke + x, random = ~1|ptt, data = combo_mod)


vec_AIC <- AIC(m.mod1.1.1.1.1.03, m.mod1.1.1.1.1.1.01, m.mod1.1.1.1.1.1.02)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2))
AICw
# sst_sd is best predictor
rm(m.mod1.1.1.1.1.1.02) # cleanup
```

# FORWARD SELECTION 8
Identify the eighth best single predictive variable for cluster placement 
``` {r MFS8}
m.mod1.1.1.1.1.1.1.01 <- mblogit(clus2 ~ species + chloro + y + lunar + n2 + eke + sst_sd + x, random = ~1|ptt, data = combo_mod)

vec_AIC <- AIC(m.mod1.1.1.1.1.1.01, m.mod1.1.1.1.1.1.1.01)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2))
AICw
# previous model was better
rm(m.mod1.1.1.1.1.1.1.01) # cleanup
```

# M.E. Including Interaction Terms
```{r ME Interactions}
# 1. species:chloro
m.mod0.1 <- mblogit(clus2 ~ species + chloro + y + lunar + n2 + eke + sst_sd + species:chloro, random = ~1|ptt, data = combo_mod)
vec_AIC <- AIC(m.mod1.1.1.1.1.1.01, m.mod0.1)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2))
AICw
# this interaction made the model better

# 2. species:y
m.mod0.2 <- mblogit(clus2 ~ species + chloro + y + lunar + n2 + eke + sst_sd + species:chloro + species:y, random = ~1|ptt, data = combo_mod)
vec_AIC <- AIC(m.mod0.1, m.mod0.2)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2))
AICw
# this interaction made the model better

# 3. species:lunar
m.mod0.3 <- mblogit(clus2 ~ species + chloro + y + lunar + n2 + eke + sst_sd + species:chloro + species:y + species:lunar, random = ~1|ptt, data = combo_mod)
vec_AIC <- AIC(m.mod0.2, m.mod0.3)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2))
AICw
# this interaction made the model better

# 4. species:n2
m.mod0.4 <- mblogit(clus2 ~ species + chloro + y + lunar + n2 + eke + sst_sd + species:chloro + species:y + species:lunar + species:n2, random = ~1|ptt, data = combo_mod)
vec_AIC <- AIC(m.mod0.3, m.mod0.4)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2))
AICw
# this interaction did not improve model

# 5. species:eke
m.mod0.4 <- mblogit(clus2 ~ species + chloro + y + lunar + n2 + eke + sst_sd + species:chloro + species:y + species:lunar + species:eke, random = ~1|ptt, data = combo_mod)
vec_AIC <- AIC(m.mod0.3, m.mod0.4)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2))
AICw
# this interaction did not improve model

# 6. species:sst_sd
m.mod0.4 <- mblogit(clus2 ~ species + chloro + y + lunar + n2 + eke + sst_sd + species:chloro + species:y + species:lunar + species:sst_sd, random = ~1|ptt, data = combo_mod)
vec_AIC <- AIC(m.mod0.3, m.mod0.4)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2))
AICw
# this interaction did not improve model

# 7. chloro:y
m.mod0.4 <- mblogit(clus2 ~ species + chloro + y + lunar + n2 + eke + sst_sd + species:chloro + species:y + species:lunar + chloro:y, random = ~1|ptt, data = combo_mod)
vec_AIC <- AIC(m.mod0.3, m.mod0.4)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2))
AICw
# this interaction did improve model

# 8. chloro:lunar
m.mod0.5 <- mblogit(clus2 ~ species + chloro + y + lunar + n2 + eke + sst_sd + species:chloro + species:y + species:lunar + chloro:y + chloro:lunar, random = ~1|ptt, data = combo_mod)
vec_AIC <- AIC(m.mod0.4, m.mod0.5)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2))
AICw
# this interaction did not improve model

# 9. chloro:n2
m.mod0.5 <- mblogit(clus2 ~ species + chloro + y + lunar + n2 + eke + sst_sd + species:chloro + species:y + species:lunar + chloro:y + chloro:n2, random = ~1|ptt, data = combo_mod)
vec_AIC <- AIC(m.mod0.4, m.mod0.5)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2))
AICw
# this interaction did not improve model

# 10. chloro:eke
m.mod0.5 <- mblogit(clus2 ~ species + chloro + y + lunar + n2 + eke + sst_sd + species:chloro + species:y + species:lunar + chloro:y + chloro:eke, random = ~1|ptt, data = combo_mod)
vec_AIC <- AIC(m.mod0.4, m.mod0.5)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2))
AICw
# this interaction did not improve model

# 11. chloro:sst_sd
m.mod0.5 <- mblogit(clus2 ~ species + chloro + y + lunar + n2 + eke + sst_sd + species:chloro + species:y + species:lunar + chloro:y + chloro:sst_sd, random = ~1|ptt, data = combo_mod)
vec_AIC <- AIC(m.mod0.4, m.mod0.5)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2))
AICw
# this interaction did not improve model\

# 12. y:lunar
m.mod0.5 <- mblogit(clus2 ~ species + chloro + y + lunar + n2 + eke + sst_sd + species:chloro + species:y + species:lunar + chloro:y + y:lunar, random = ~1|ptt, data = combo_mod)
vec_AIC <- AIC(m.mod0.4, m.mod0.5)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2))
AICw
# this interaction did improve model

# 13. y:n2
m.mod0.6 <- mblogit(clus2 ~ species + chloro + y + lunar + n2 + eke + sst_sd + species:chloro + species:y + species:lunar + chloro:y + y:lunar + y:n2, random = ~1|ptt, data = combo_mod)
vec_AIC <- AIC(m.mod0.5, m.mod0.6)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2))
AICw
# this interaction did improve model

# 14. y:eke
m.mod0.7 <- mblogit(clus2 ~ species + chloro + y + lunar + n2 + eke + sst_sd + species:chloro + species:y + species:lunar + chloro:y + y:lunar + y:n2 + y:eke, random = ~1|ptt, data = combo_mod)
vec_AIC <- AIC(m.mod0.6, m.mod0.7)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2))
AICw
# this interaction did not improve model

# 15. y:sst_sd
m.mod0.7 <- mblogit(clus2 ~ species + chloro + y + lunar + n2 + eke + sst_sd + species:chloro + species:y + species:lunar + chloro:y + y:lunar + y:n2 + y:sst_sd, random = ~1|ptt, data = combo_mod)
vec_AIC <- AIC(m.mod0.6, m.mod0.7)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2))
AICw
# this interaction did not improve model

# 16. lunar:n2
m.mod0.7 <- mblogit(clus2 ~ species + chloro + y + lunar + n2 + eke + sst_sd + species:chloro + species:y + species:lunar + chloro:y + y:lunar + y:n2 + lunar:n2, random = ~1|ptt, data = combo_mod)
vec_AIC <- AIC(m.mod0.6, m.mod0.7)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2))
AICw
# this interaction did not improve model

# 17. lunar:eke
m.mod0.7 <- mblogit(clus2 ~ species + chloro + y + lunar + n2 + eke + sst_sd + species:chloro + species:y + species:lunar + chloro:y + y:lunar + y:n2 + lunar:eke, random = ~1|ptt, data = combo_mod)
vec_AIC <- AIC(m.mod0.6, m.mod0.7)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2))
AICw
# this interaction did not improve model

# 18. lunar:sst_sd
m.mod0.7 <- mblogit(clus2 ~ species + chloro + y + lunar + n2 + eke + sst_sd + species:chloro + species:y + species:lunar + chloro:y + y:lunar + y:n2 + lunar:sst_sd, random = ~1|ptt, data = combo_mod)
vec_AIC <- AIC(m.mod0.6, m.mod0.7)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2))
AICw
# this interaction did not improve model

# 19. n2:eke
m.mod0.7 <- mblogit(clus2 ~ species + chloro + y + lunar + n2 + eke + sst_sd + species:chloro + species:y + species:lunar + chloro:y + y:lunar + y:n2 + n2:eke, random = ~1|ptt, data = combo_mod)
vec_AIC <- AIC(m.mod0.6, m.mod0.7)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2))
AICw
# this interaction did improve model

# 20. n2:sst_sd
m.mod0.8 <- mblogit(clus2 ~ species + chloro + y + lunar + n2 + eke + sst_sd + species:chloro + species:y + species:lunar + chloro:y + y:lunar + y:n2 + n2:eke + n2:sst_sd, random = ~1|ptt, data = combo_mod)
vec_AIC <- AIC(m.mod0.7, m.mod0.8)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2))
AICw
# this interaction did not improve model

# 21. eke:sst_sd
m.mod0.9 <- mblogit(clus2 ~ species + chloro + y + lunar + n2 + eke + sst_sd + species:chloro + species:y + species:lunar + chloro:y + y:lunar + y:n2 + n2:eke + eke:sst_sd, random = ~1|ptt, data = combo_mod)
vec_AIC <- AIC(m.mod0.7, m.mod0.9)[,2]
dAIC <- vec_AIC - min(vec_AIC)
AICw <- exp(-dAIC/2) / sum(exp(-dAIC/2))
AICw
# this interaction did not improve model
```

# The mixed effet multinomial logistic regression model predicting blue and mako shark cluster placement is: m.mod0.7 

species + chloro + y + lunar + n2 + eke + sst_sd + species:chloro + species:y + species:lunar + chloro:y + y:lunar + y:n2 + n2:eke, random = ~1|ptt

Species
Primary Productivity
Latitude
Lunar
n2
Eddy Kinetic Energy
Sea-surface temperature standard deviation

Species : Primary Productivity
Species : Latitude
Species : Lunar Illumination
Primary Productivity : Latitude
Latitude : Lunar Illumination
Latitude : Buoyancy Frequency
Buoyancy Frequency : Eddy Kinetic Energy


# Examining Model Results
Lets consider how the probability of each cluster changes with each individual predictor while holding the other model variables at their mean values
```{r chloro}
# Chlorophyll
p_chloro <- data.frame(
  chloro = rep(seq(0,6.5, 0.1), times = 8), 
  species = as.factor(rep(c("I. oxyrinchus", "P. glauca"), each = 264)),
  y = rep(c(15, 45, 15, 45), each = 132),
  lunar = as.factor(rep(c(1,2,1,2), times = 132)),
  n2 = rep(c(6.427669e-05), times = 528),
  sst_sd = rep(c(0.2973896), times = 528),
  eke = rep( c(-1.315512), times = 528),
  ptt = as.factor(rep(c("78683", "106754"), each = 264))
  )

p_chloro <- cbind(p_chloro, predict(m.mod0.7, newdata = p_chloro, type = "response", se.fit = FALSE))
lchloro <- melt(p_chloro, id.vars = c("species", "chloro", "y", "lunar", "n2", "sst_sd", "eke", "ptt"), measure.vars = c("3", "1", "2", "4", "5", "6", "7"))
lchloro$y <- factor(lchloro$y) 

lchloro <- filter(lchloro, lunar == 1)

ggplot(data = lchloro) +
  geom_line(aes(x = chloro, y = value, color = variable, linetype = y)) +
  facet_grid(species ~ y)
```

```{r n2}
# Buoyancy Frequency
p_n2 <- data.frame(
  chloro = rep(c(0.5905629), times = 1368), 
  species = as.factor(rep(c("I. oxyrinchus", "P. glauca"), each = 684)),
  y = rep(c(36.74643), times = 1368),
  lunar = as.factor(rep(c(1,2,1,2), times = 342)),
  n2 = rep(seq(-4e-06, 0.00085, 0.000005), times = 8),
  sst_sd = rep(c(0.2973896), times = 1368),
  eke = rep( c(-1.747338, -0.7923545, -1.747338, -0.7923545), each = 342),
  ptt = as.factor(rep(c("78683", "106754"), each = 684))
  )

p_n2 <- cbind(p_n2, predict(m.mod0.7, newdata = p_n2, type = "response", se.fit = FALSE))
ln2 <- melt(p_n2, id.vars = c("species", "chloro", "y", "lunar", "n2", "sst_sd", "eke", "ptt"), measure.vars = c("3", "1", "2", "4", "5", "6", "7"))
ln2$eke <- factor(ln2$eke)
ln2 <- filter(ln2, lunar == 1)
ln2$variable <- factor(ln2$variable, levels = c(1,2,3,4,5,6,7))

ggplot(data = ln2) +
  geom_line(aes(x = n2, y = value, color = variable, linetype = eke)) +
  scale_x_continuous(limits = c(-4e-06, 0.00085)) +
  facet_grid(species ~ eke)

```

# Cluster Conditions
This chunk computes the average environmental conditions within each cluster
```{r cluster conditions}
combo_mod$Month <- as.numeric(combo_mod$Month)
combo_mod$Year <- as.numeric(combo_mod$Year)
combo_mod$lunar <- as.numeric(combo_mod$lunar)

combo_mod %>% group_by(cluster) %>%
  summarize(
    sst = mean(sst, na.rm = TRUE),
    sst_sd = mean(sst_sd, na.rm = TRUE),
    ssh = mean(ssh, na.rm = TRUE),
    ssh_sd = mean(ssh_sd, na.rm = TRUE),
    eke = mean(eke, na.rm = TRUE),
    ild.5 = mean(ild.5, na.rm = TRUE),
    n2 = mean(n2, na.rm = TRUE),
    lunar = median(lunar, na.rm = TRUE),
    bathy = mean(bathy, na.rm = TRUE),
    chloro = mean(chloro, na.rm = TRUE),
    temp200 = mean(Temp200, na.rm = TRUE),
    month = median(Month, na.rm = TRUE),
    year = median(Year, na.rm = TRUE)
  )
```

1. create boxplots of conditions within each cluster
```{r conditon boxplots}
# sst:
ggplot() +
  geom_boxplot(data = omega_combo, aes(y = sst, color = cluster)) +
  scale_x_continuous(breaks = c(), labels = c()) +
  theme_classic()

# sst_sd:
ggplot() +
  geom_boxplot(data = omega_combo, aes(y = sst_sd, color = cluster)) +
  scale_x_continuous(breaks = c(), labels = c()) +
  theme_classic()

# ssh:
ggplot() +
  geom_boxplot(data = omega_combo, aes(y = ssh, color = cluster)) +
  scale_x_continuous(breaks = c(), labels = c()) +
  theme_classic()

# ssh_sd:
ggplot() +
  geom_boxplot(data = omega_combo, aes(y = ssh_sd, color = cluster)) +
  scale_x_continuous(breaks = c(), labels = c()) +
  theme_classic()

# eke:
ggplot() +
  geom_boxplot(data = omega_combo, aes(y = eke, color = cluster)) +
  scale_x_continuous(breaks = c(), labels = c()) +
  theme_classic()

# ild.5:
ggplot() +
  geom_boxplot(data = omega_combo, aes(y = ild.5, color = cluster)) +
  scale_x_continuous(breaks = c(), labels = c()) +
  scale_y_reverse()+
  theme_classic()

# n2:
ggplot() +
  geom_boxplot(data = omega_combo, aes(y = n2, color = cluster)) +
  scale_x_continuous(breaks = c(), labels = c()) +
  theme_classic()

# bathy:
ggplot() +
  geom_boxplot(data = omega_combo, aes(y = bathy, color = cluster)) +
  scale_x_continuous(breaks = c(), labels = c()) +
  theme_classic()

# chloro:
ggplot() +
  geom_boxplot(data = omega_combo, aes(y = chloro, color = cluster)) +
  scale_x_continuous(breaks = c(), labels = c()) +
  theme_classic()

# Temp200:
ggplot() +
  geom_boxplot(data = combo_mod, aes(y = Temp200, color = cluster)) +
  scale_x_continuous(breaks = c(), labels = c()) +
  theme_classic()
```

2. When did clusters occur:
Is there a pattern in temporal occurance of different clusters?
```{r when bar plots}
# what month is most common in the data overall
c0 <- omega_combo 
c0$cluster <- 0

ggplot() +
  geom_bar(data = c0, aes(x = Month)) +
  geom_bar(data = omega_combo, aes(x = Month, color = cluster, fill = cluster)) +
  facet_wrap(~cluster) +
  theme_classic()

```

3. Who performed each cluster
What was the frequency at which individual sharks engaged in each cluster
```{r who bar blots}
ggplot() +
  geom_bar(data = c0, aes(x = ptt)) +
  geom_bar(data = omega_combo, aes(x = ptt, fill = cluster)) +
  facet_wrap(~species) +
  coord_flip() +
  theme_classic()
```


## NOTES FROM 2022/03 ###
After consulting with the MPG and with Camrin, it seems that there are really only two broad behavioral states that blue and mako sharks exhibit in this dataset. These are epipelagic surface use and diel-vertical migration. However, there appears to be considerable variation in fine-scale patterns of movement within each of these two behavior classes

To examine both coarse and fine scale behavior patterns, we have decided to try a nested clustering approach. Using the course-resolution TAD summaries from the first analysis itteration, we will perform clustering aimed at identifying tracking days from the two broad behavior types. Next, using fine-scale TAD summaries reconstructed from a subset of tracking days for which time-series data are available, we will perform a second round of clustering in an attempt to identify fine-scale behavior variations. 

For this analysis, I will exclude any tracking days which occur in water more shallow than 1000 meters. This is because the environmental drivers of behavior may be confounded with physical habitat constrictions in these locations. For example, conditions on the continental shelf attract blue sharks in the spring and summer to regions where they cannot utilize their normal vertical range - therefore epipelagic surface-use behavior may appear to be more common in warmer waters with high productivity simply because it is limited to shelf occurrences. In reality, the drivers of this behavior in the pelagic habitat of blue sharks may be much different.

# Course Clustering Data
This chunk compiles the data for a course-resolution cluster anlaysis of blue and mako shark behavior:
```{r course cluster: data}
# filter tracking days in water shallower than 1000 meters
course_tad <- filter(omega_combo, bathy <= -1000)
course_m7 <- filter(m7, depth <= -1000)

# select relevant columns
course_tad <- course_tad %>% dplyr::select(c("Date", "bin1", "bin2", "bin3", "bin4" , "bin5", "bin6", "bin7", "bin8", "ptt", "species"))
course_m7 <- course_m7 %>% dplyr::select(c("Date", "bin1", "bin2", "bin3", "bin4" , "bin5", "bin6", "bin7", "bin8", "ptt", "species"))

# add lucky mako 7 to the main dataset
course_m7$ptt <- factor(course_m7$ptt)
course_tad <- rbind(course_tad, course_m7) # results in 1337 tracking days for clustering
rm(course_m7)

# adding kode back to course_tad
course_tad <- course_tad %>% mutate(kode = paste(Date, ptt, sep = "_"))
course_blue <- course_tad %>% filter(species == "P. glauca") # 824 tracking days
course_mako <- course_tad %>% filter(species == "I. oxyrinchus") # 513 tracking days
```

This chunk performs cluster analysis on the course-resolution data:
```{r course cluster: analysis}
## heirarchical clustering code
Ctad <-  course_tad %>% dplyr::select(c("bin1", "bin2", "bin3", "bin4" , "bin5", "bin6", "bin7", "bin8"))
Ctad[Ctad < 10] <- 0 # set low occurrence bins to 0 - Sal's recommendation but may need to be omitted

combo_dist <- dist(Ctad, method = "manhattan")

Clust <- hclust(combo_dist, method = "average")

## statistical assessments for optimum number of clusters
NbClust(data = Ctad, diss = NULL, distance = "manhattan", min.nc = 2,
        max.nc = 20, method = "average", index = "ch")
# LowToZero: 12 clusters
# Raw: 8 clusters

NbClust(data = Ctad, diss = NULL, distance = "manhattan", min.nc = 2,
        max.nc = 20, method = "average", index = "duda")
# LowToZero: 2 clusters
# Raw: 2 clusters

NbClust(data = Ctad, diss = NULL, distance = "manhattan", min.nc = 2,
        max.nc = 20, method = "average", index = "cindex")
# LowToZero: 20 clusters 
# Raw: 20 clusters

NbClust(data = Ctad, diss = NULL, distance = "manhattan", min.nc = 2,
        max.nc = 20, method = "average", index = "beale")
# LowToZero: 2 clusters
# Raw: 2 clusters

## examining cluster results
combo_sub_grp <- cutree(Clust, k = 4)
table(combo_sub_grp)
# NOTE: Setting low bins to 0 seems to affect the NbClust outcomes but hardly affects the clustering scheme itself based on the cuttree outcomes for k = 1:5
Ctad$cluster <- combo_sub_grp

course_tad$cluster <- combo_sub_grp

clust_stamp <- course_tad %>% dplyr::select(kode, species, cluster)
```

This next chunk visualizes the different course-resolution clusters created on line 2962-2969 of the above chunk:
```{r course cluster: visualization}
combo_series <- rbind(mako_series, blue_series)
combo_series$Local_Time <- as_hms(combo_series$Local_Time)
course_series <- inner_join(combo_series, clust_stamp, by = "kode")
unique(course_series$kode)

# daily cluster kernal density heatmaps
ggplot(data = course_series, aes(x = Local_Time, y = depth) ) +
  stat_density_2d(aes(fill = ((..ndensity..))), geom = "raster", contour = FALSE) +
  facet_wrap(~cluster) +
  scale_fill_viridis(option = "magma") + 
  labs(fill = "Density") +
  scale_y_reverse() +
  scale_x_continuous(breaks = c(0, 43200, 86400), labels = c("00:00:00", "12:00:00", "23:00:00")) +
  xlab("Time of Day") +
  ylab("Depth(m)") +
  theme_minimal()

# daily cluster point maps
ggplot(data = course_series) +
  geom_point(aes(x = Local_Time, y = depth, color = species), alpha = 0.5) +
  facet_grid(cluster ~ species) + 
  scale_y_reverse() +
  theme_classic()
```




